### 2017-06-28 literaturereview.Rmd
Literature Review

---
title: "“Question Quality in Community Q&A” Notes"
author: "Lisa Oshita"
date: "June 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Question Content
* length: word count and sentence count
* text: ngram features from title and body

###Global/local topic models
* global:
  + topic information considered as another predictor of question quality
  + reflects what the entire question is about
* local:
  + captures internal structural aspects of questions 
  + Sentence topic model: capture latent aspects at internal level

###Results
* some correlation between view count and question quality 
* content (length and text) gave signficant improvement in classification accuracy (over ViewCount)- indicates that words used capture quality 
* good questions tended to be shorter 
* prediction accuracy increased with global topics- latent topics can be predictive of quality
* content + local topic models = better performance over content + global - internal structures of a question indicate quality

###Discussion
* good questions: 
* aren't repeats (user has done research)
* receive more answers on average (5 compared to 2.5), also receive longer answers and more comments
* "live longer"
* "latent topical aspects shared between related questions are good predictors of question quality"

##Ideas for research
* consider question and title length (predict that concise questions will get answered quicker than lengthy ones)
* extract topic information from title and body of the question?