### 2017-06-29 new_variables.Rmd
creating response variable

---
title: "New Variables"
author: "Lisa Oshita"
date: "June 29, 2017"
output: html_document
---

##Creating time to event variable
```{r}
out$time_until_answer <- out$first_answer_date - out$post_date
```