### 2017-06-30 new_variables.Rmd
practicing with stringr functions

---
title: "New Variables"
author: "Lisa Oshita"
date: "June 29, 2017"
output:
  pdf_document: default
  html_document: default
---

### Creating time to event variable
observation 15 has a negative value?
```{r}

out$time_until_answer <- out$first_answer_date - out$post_date

empty <- which(is.na(out$time_until_answer))
for (i in empty) {
  out$time_until_answer[i] <- out$download_date[i] - out$post_date[i]
}

```

### Practicing with stringr 
```{r}
library(stringr)
library(rebus)

#variable indicating if title contains questionmark 
out$contain_questionmark <- str_detect(out$title, pattern = QUESTION)
```

### Practicing with product categorization 
```{r}
# having problems with regex function, not giving same output that Data camp showed 
x <- c("Cat", "CAT", "cAt")
str_view(x, regex("cat", ignore_case = TRUE), match = TRUE)

# not working
phones <- regex("phone", ignore_case = TRUE)
out$product_type[str_detect(out$device, pattern = phones)] <- "Phone"
```