### 2017-07-07 practice.Rmd
practice with text mining

---
title: "Practicing with the data"
author: "Lisa Oshita"
date: "June 29, 2017"
output:
  pdf_document: default
  html_document: default
---
```{r}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))
```

### Creating time to event variable
censoring indicator variable: answered
```{r}
out$time_until_answer <- out$first_answer_date - out$post_date
#filling in the NAs
empty <- which(is.na(out$time_until_answer))
for (i in empty) {
  out$time_until_answer[i] <- out$download_date[i] - out$post_date[i]
}

#26 observations in out dataframe have negative time to event variable
length(which(out$time_until_answer < 0))
library(dplyr)
out %>% 
  filter(time_until_answer < 0) %>%
  select(post_date, first_answer_date, answered)
```

### Practicing with stringr 
```{r}
library(stringr)
library(rebus)

#variable indicating if title contains questionmark 
out$contain_questionmark <- str_detect(out$title, pattern = QUESTION)

#practicing with product categorization
phones <- regex("phone", ignore_case = TRUE)
out$product_type[str_detect(out$device, pattern = phones)] <- "Phone"
```

### Practicing with dplyr
```{r}
library(stringr)
library(dplyr)

#using only observations in English
out_english <- out %>% 
                tbl_df() %>% 
                filter(langid == "en")

#view average time until answered grouped by each category 
out_english%>%
  group_by(category) %>%
  summarise(avg_time = mean(time_until_answer, na.rm = TRUE), num = n()) %>%
  arrange(avg_time)

#view average time until answered grouped by whether or not question contains "?", for each category
#seems like for a lot of categories, questions with question marks get answered faster
out_english$contain_questionmark <- str_detect(out_english$title, pattern = QUESTION)
out_english %>%
  group_by(category, contain_questionmark) %>%
  summarise(avg_time = mean(time_until_answer, na.rm = TRUE), num = n()) %>%
  arrange(category, avg_time)

#looking at average time until answered for intervals of title length
#creating title length intervals
title_length <- str_length(out_english$title)
out_english$title_length_intervals <- cut(title_length, breaks = 8)
#don't really see anything interesting here 
out_english %>%
  group_by(title_length_intervals) %>%
  summarise(avg_time = mean(time_until_answer, na.rm = TRUE), num = n())

```

### practicing with ggplot
```{r}
library(ggplot2)
library(stringr)

#using only observations with a positive time_until_answer value (26 observations have negative time until answer values)
out1_positive <- out_english %>% filter(time_until_answer > 0)

#histogram of time_until_answer
ggplot(out1_positive, aes(x = time_until_answer)) + 
  geom_histogram(bins = 10)

#plot of number of tags and time_until_answer
ggplot(out1_positive, aes(x = factor(n_tags), y = time_until_answer)) + 
  geom_point(position = "jitter")

#plot with title length 
out1_positive$title_length <- str_length(out1_positive$title)

ggplot(out1_positive, aes(x = title_length, y = time_until_answer)) + 
  geom_point()

#looking at only data for phones category 
phones_only <- out1_positive %>% filter(category == "Phone")
ggplot(phones_only, aes(x = title_length, y = time_until_answer)) + 
  geom_point(size = 3, alpha = 0.5)
```

### practicing with text mining
```{r}
library(dplyr)
#working only with the 2000 questions that were answered the fastest
arranged <- out1_positive %>% arrange(time_until_answer)
fastest <- arranged[1:2000,]

library(qdap)
library(tm)
# INSTALLING RWEKA NOT WORKING
# library(RWeka)
# 
# titles <- fastest$title
# 
# #clean titles 
# qdap_clean <- function(x) {
#   x <- replace_abbreviation(x)
#   x <- replace_contraction(x)
#   x <- replace_number(x)
#   x <- replace_ordinal(x)
#   x <- replace_symbol(x)
#   x <- tolower(x)
#   return(x)
# }
# tm_clean <- function(corpus) {
#   tm_clean <- tm_map(corpus, removePunctuation)
#   corpus <- tm_map(corpus, stripWhitespace)
#   corpus <- tm_map(corpus, removeWords, stopwords("en"))
#   return(corpus)
# }
# 
# tokenizer <- function(x) {
#   NGramTokenizer(x, Weka_control(min = 2, max = 3)) }
# 
# titles <- qdap_clean(titles)
# titles_cor <- VCorpus(VectorSource(titles))
# titles_corpus <- tm_clean(titles_cor)
# 
# titles_tdm <- TermDocumentMatrix(titles_corpus, )
```