### 2017-07-11 practice_parametricdist.Rmd
fitting parametric distributions to the data

---
title: "Practice with parametric distributions"
author: "Lisa Oshita"
date: "July 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### import the data
```{r}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

#creating time to event variable
out$time_until_answer <- out$first_answer_date - out$post_date
#filling in the NAs
empty <- which(is.na(out$time_until_answer))
for (i in empty) {
  out$time_until_answer[i] <- out$download_date[i] - out$post_date[i]
}

#subsetting to include only observations in english with positive time to event values
library(dplyr)
out1 <- out %>% 
          tbl_df() %>% 
          filter(langid == "en") %>%
          filter(time_until_answer > 0)
```

### finding parameter estimates with flexsurv package
```{r}
library(flexsurv)

#weibull distribution
weibull <- flexsurvreg(surv_object~1, data = out1, dist = "weibull")
weibull
plot(weibull)
#shape = 0.228, scale = 1320000, AIC = 123791.7
qqplot(rweibull(1e+5, shape = 0.228, scale = 1320000), out1$time_until_answer)

#exponential distribution
exp <- flexsurvreg(surv_object~1, data = out1, dist = "exp")
exp
plot(exp)
#rate = 4.98e-07, AIC = 153610

#lognormal distribution
lnormal <- flexsurvreg(surv_object~1, data = out1, dist = "lnorm")
lnormal
plot(lnormal)
#meanlog = 12.05, sdlog = 4.86, AIC = 122112.1

#logistic distribution
logistic <- flexsurvreg(surv_object~1, data = out1, dist = "llogis")
logistic
plot(logistic)
#shape = 0.332, scale = 119000, AIC = 122482
```

### setting up probability plots to assess fit
```{r}
# library(car) #for the qqPlot function
# par(mfrow = c(2,2))
# qqPlot(out1$time_until_answer, distribution = "weibull", param.list = list(shape = 0.228, scale = 1320000), plot.type = "Q-Q")
qqplot(rweibull(n = 8000, shape = 0.228, scale = 1320000), out1$time_until_answer)
abline(0,1)
qqplot(rlnorm(n = 8000, meanlog = 12.05, sdlog = 4.86), out1$time_until_answer)
abline(0,1)
#are the plots actually following the line?- maybe scale too large to see it correctly 
```

### learning about flexsurvreg 
```{r}
time <- c(5,8,12,24,32,17,16,17,19,30)
censor <- c(1,1,1,1,1,1,0,0,0,0)
example <- data.frame(time, censor)

s <- Surv(example$time, example$censor, type = "right")
exponential <- flexsurvreg(s~1, data = example, dist = "exp")
exponential
survreg(s~1, data = example, dist = "exp")

data(ovarian)
fitg <- flexsurvreg(formula = Surv(futime, fustat) ~ 1, data = ovarian, dist="gengamma")
fitg
fitw <- flexsurvreg(formula = Surv(futime, fustat) ~ 1, data = ovarian, dist="weibull")
fitw
plot(fitg)
lines(fitw, col="blue", lwd.ci=1, lty.ci=1)
```

### parametric distributions using fitdistrplus package NOT WORKINGG
```{r}
library(fitdistrplus)

#creating the data frame to use in the function
left <- right <- out1$time_until_answer
right[out1$answered == 0] <- NA_integer_
censdata <- data.frame(left, right)

#fitweibull <- fitdistcens(censdata, distr = "weibull")

```

### practicing with fitdistcens function 
```{r}
data("salinity")
str(salinity)
data("smokedfish")

fsal.ln <- fitdistcens(salinity, "lnorm")
fsal.ln

log10EC50 <-log10(fluazinam)
fln <- fitdistcens(log10EC50,"norm")
fln
summary(fln)
plot(fln)
```