### 2017-07-12 Hours.Rmd
7/11 hours

> Summer Research Hours Log

Date   | Hours | Time        | Description
------ | ----- | ----------- | -----------------------------------------
6/26/2017 | 1.5 | 2:30-4:00PM | Integrated R and Github, familiarized myself with GitHub
6/27/2017 | 1 | 11:00-12:00PM | Explored the Q&A forum, read and took notes on part of the study posted on GitHub
6/28/2017 | 3.5 | 11:30-3:00PM | Finished reading and taking notes on the study, learned how to upload to Github through R studio, started the stringr data camp course
6/28/2017 | 5 | 5:00-10:00PM | Continued with data camp courses
6/29/2017 | 3 | 7:00-10:00AM | Continued with data camp courses, brainstormed variables to consider in the model
6/29/2017 | 3.5 | 3:00-6:30PM | Responded to issues, redid stringr course and created a cheatsheet
6/30/2017 | 3.25 | 8:15-11:30AM | Redid/completed stringr course, practiced with stringr functions on the data set
6/30/2017 | 3 | 3:30-6:30PM | Ran into problems with stringr functions and the dataset- tried to get them to work, moved onto the next data camp course
6/30/2017 | 2 | 6:30-9:30PM | Writing functions data camp course + cheatsheet
7/1/2017 | 2 | 8:30-10:30AM | Completed writing functions course, started dplyr course 
7/2/2017 | 1 | 9:00-10:00PM | Continued working on the dplyr course + cheatsheet
7/3/2017 | 2.25 | 10:00-12:15PM | Finished dplyr course
7/3/2017 | 1.5 | 4:00-5:30PM | Practiced applying dplyr functions to the dataset, started on ggplot2 course
7/3/2017 | 2 | 6:30-8:30PM | Continued working on ggplot2 course + cheatsheet
7/4/2017 | 1.5 | 2:00-3:30PM | Continued with ggplot2 course
7/5/2017 | 3 | 9:30-12:30PM | Finished ggplot2 course, practiced using ggplot on the dataset, reviewed STAT 417 notes
7/5/2017 | 2.25 | 2:45-5:00PM | Started the text mining course + cheatsheet
7/6/2017 | 1.5 | 7:30-9:00AM | Continued with text mining course
7/6/2017 | 2.75 | 2:15-5:00PM | Finished text mining course + cheatsheet
7/6/2017 | 2.5 | 6:30-9:00PM | Tried to apply text mining techniques to the dataset, ran into RWeka package, uploaded hours timesheet to Github
7/7/2017 | 4.75 | 11:15-5:00PM | Learned and practiced with functions in the survival package on the data set, came up with descriptive summaries, explored the response variable 
7/10/2017 | 1 | 9:00-10:00AM | Started learning how to fit parametric distributions in r
7/10/2017 | 4 | 1:00- 5:00PM | Experimented with calculating MLEs from the data for different parametric distributions, tried to create probability plots to assess the fits for each distribution
7/10/2017 | 3 | 6:30-9:30PM | Continued working with parametric distributions, experimented with fitdistrplus package to fit parametric distributions
7/11/2017 | 4.25 | 10:45-3:00PM | Tried to get the probability plots to work, used techniques learned in data camp courses and survival analysis to explore the data set and get an idea of what variables may be important 
7/11/2017 | 3.75 | 6:00-9:45PM | Continued to explore the data set

>> __Total Hours: 68.75__