### 2017-07-12 data_exploration.Rmd
summarizing findings from variable exploration and from using survival analysis techniques

---
title: "Descriptive Statistics"
author: "Lisa Oshita"
date: "July 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic descriptive statistics/exploring the variables
### using out_english data set: 
* about 8000 observations
* around 64% of questions were answered (~ 36% unanswered/censored)

### examining distribution of time_until_answer
* min = 30 sec, max = 7772000 sec (~90 days) (id = 389776)
* mean = 1281000 sec (~15 days), median = 31650 sec (~8 hours)
* 1st quartile = 3134 sec (52.2 min), 3rd quartile = 1975000 (~23 days) (25% of the questions wre answered before 52 min, and 75% of the questions were answered after 23 days, middle 50% of the data falls between 52 min and 23 days)
* distibution of response is extremely skewed right 
* from distribution it seems like quite a bit of data falls between 30 sec and 1200000 (~14 days)

### examining time_until_answer by categories 
* phone, mac, tablet, pc, and electronics have the most questions (out of 15 total categories)
* Skills category has shortest average time_until_answer, but only has 5 in that category
* Mac has the second shortest average time_until_answer, with a good amount of observations (666)
* PC and then Phone have next shortest average times, with 413 and 2148 observations 
* Camera has the longest average time, with 269 observations
* looking at categories according to median time_until_answer gives similar results (skills, mac and phone have shortest median times)
* distributions of time_until_answer for Mac, PC, Phone are extremely skewed right (exactly like the distribution of time_until_answer for all categories)

### examining time_until_answer according to whether or not it contains a question mark
* 1193/7760 (15%) of the questions contain a question mark
* for almost all categories, questions that contain a question mark seem to get answered faster (lower average and median times) 
* but less than half of the total questions for each category contain question marks- may be contributing to the lower values 
* around 71% of the questions that contain question marks were answered 
* questions about Mac, PC, and skills (but only contains 5 obs), have the highest proportion of questions that contain "?"

### examining title length (look at length of text and device title as well)
* mean = 42 characters
* median = 40 characters
* longest title is 89 characters, shortest is 12 characters
* seems like questions with longer titles might have shorter average time_until_answer
* Mac has average title length of 44 characters, average time_until_answer = 453980.6 seconds (which is on the shorter end of values), while those with shorter titles seem to have longer times

### examining text length
* shortest text: 20 characters, longest text: 4814.0
* from histogram of text lengths, a bulk of the questions have text lengths between 20 and 1000 characters long (distribution skewed right)
* seems like categories with longer median text lengths have shorter time_until_answer values

### device length 
* min = 0, max = 136 (id = 390271), mean = 17, median = 16
* distribution of device lengths is skewed right
* didn't really see anything interesting here

### number of tags included in a question
* max number of tags = 4
* 56% of questions don't include any tags
* 43% of the questions that do have a tag only have 1 tag
* seems like there MAY be a slight tendency for questions with more tags to get answers quicker (shorter avg time_until_answer, but also have not as many observations)

### number of images included
* only 4% of the questions contain any images
* min = 0, max = 13 (id = 396580)
* there MAY be some evidence that questions that include images seems to get answers quicker

## Survival analysis statistics 
### KM estimator/curve
* from the KM curve, the survival probability drops drastically from 1 down to about 0.4 before 400,000 seconds (~5 days). After that point, survival curve seems to decrease slowly and remain around 0.3-0.4 (due to the largest times being censored) 
* lowest survival probability is 0.346

### quantiles
* (median) At 30,158 seconds (8.38 hours), 50% of the questions have been answered
* at 857 seconds (14.3 min), 10% of the questions have been answered
* at 3163 seconds (52.7 min), 25% of the questions have been answered
* there is no time for which 75th percentile (since the lowest survival probability = 0.346, and it never reaches 0) (means that we'll never get a total of 75% of questions answered, there will always be some questions that "survive")

### hazard rates
* instantaneous hazard drops pretty drastically after question is first posted, before 400,000 seconds (~4-5 days), it appears that hazard rate is really close to 0
* looking at the plot, there's a spike in instantaneous hazard around 200,000 seconds (~2.3 days)
* cumulative hazard rate: increases very quickly until about 400000 seconds (~4-5 days) (indicating that hazard is increasing on that interval), when it starts to even out until it coasts at 1 (indicating that hazard is decreasing
* do the Nelson Aalen and KM estimators of cumulative hazard look the same becasue the data set is so large?

#at 30,158 seconds (502.6 min, 8.38 hours), 50% of the questions have been answered
#mean survival time = 2,748,083 (45801.38 min, 763.3563hrs, 31.8 days)