### 2017-07-22 functions.Rmd

---
title: "Functions Created"
author: "Lisa Oshita"
date: "July 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setup <- function() {
  dir <- file.path(getwd(),"data")
  out <- read.csv(file.path(dir, "answers_data.csv"))

  library(dplyr)
  out_english <- out %>% 
                  tbl_df() %>%
                  filter(langid == "en")

  out_english$time_until_answer <- (out_english$first_answer_date - out_english$post_date)/3600
  empty <- which(is.na(out_english$time_until_answer))
  for (i in empty) {
    out_english$time_until_answer[i] <- (out_english$download_date[i] - out_english$post_date[i])/3600
}
}

```

```{r}

plot_surv <- function(survfit, data) {

  library(ggfortify)
  library(directlabels)

  df <- fortify(survfit, data = data)

  if (!("strata" %in% names(df))) {
    ggplot(df, aes(x = time, y = surv)) + 
    geom_line() + 
    scale_y_continuous("Survival Probability") +
    scale_x_continuous("Time")
    ggtitle("Survival Curve")
  }

  if ("strata" %in% names(df)) {
    ggplot(df, aes(x = time, y = surv, color = strata)) + 
      geom_line() + 
      scale_y_continuous("Survival Probabilities") +
      scale_x_continuous("Time") + 
      ggtitle("Survival Curves") + 
      geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
  }
}
```