### 2017-07-28 functions.Rmd
finished work with tags, updated hours, tweaked functions/other files

---
title: "Functions Created"
author: "Lisa Oshita"
date: "July 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggfortify)
library(directlabels)
library(tm)
library(qdap)
```

```{r}
#imports data set, includes only questions in English, creates time_until_answer (in hrs)
setup <- function() {
  dir <- file.path(getwd(),"data")
  out <- read.csv(file.path(dir, "answers_data.csv"))

  out_english <- out %>% 
                  tbl_df() %>%
                  filter(langid == "en")

  out_english$time_until_answer <- (out_english$first_answer_date - out_english$post_date)/3600
  empty <- which(is.na(out_english$time_until_answer))
  for (i in empty) {
    out_english$time_until_answer[i] <- (out_english$download_date[i] - out_english$post_date[i])/3600
}
}
```

```{r}
#function to plot survfit objects
#input: survfit object, data, optional x-axis limits
plot_surv <- function(survfit, data, xlim) {

  df <- fortify(survfit, data = data)

  if (missing(xlim)) {

    if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time")
          ggtitle("Survival Curve")
        }
    if ("strata" %in% names(df)) {
      ggplot(df, aes(x = time, y = surv, color = strata)) + 
        geom_line() + 
        scale_y_continuous("Survival Probabilities") +
        scale_x_continuous("Time") + 
        ggtitle("Survival Curves") + 
        geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
    }
  } else {

      if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time", limits = xlim)
          ggtitle("Survival Curve")
      }

      if ("strata" %in% names(df)) {
        ggplot(df, aes(x = time, y = surv, color = strata)) + 
          geom_line() + 
          scale_y_continuous("Survival Probabilities") +
          scale_x_continuous("Time", limits = xlim) + 
          ggtitle("Survival Curves") + 
          geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
      }
  }
}
```

```{r}
#function to clean corpus
clean_corpus <- function(corpus){

  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, content_transformer(tolower))
  corpus <- tm_map(corpus, removeWords, c(stopwords("en"), "can", "will", "cant", "wont", "works", "get", "help", "need"))
  return(corpus)
}

#returns a data frame of words and frequencies sorted by most to least frequent
#input: character vector of words, n is optional (to output top n most frequent words)
freq_terms <- function(vec, n) {

  source <- VectorSource(vec)
  corpus <- VCorpus(source)
  cleaned_corpus <- clean_corpus(corpus) 
  dtm <- DocumentTermMatrix(cleaned_corpus)
  m <- as.matrix(dtm) 
  freq <- colSums(m)
  freq <- sort(freq, decreasing = TRUE)
  freq_df <- data.frame(word = names(freq), frequency = unname(freq))

  if (missing(n)) {
    return(freq_df)
  }
  else {
    return(freq_df[1:n,])
  }
}
```