### 2017-07-31 freq_words.Rmd
analyzed text, added to other files, updated hours

---
title: "Frequent Words"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* **Idea: there are words that show up frequently among answered questions and unanswered questions, which might be able to be used as predictors in the model**

```{r, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

library(dplyr)
x <- out %>% 
  tbl_df() %>%
  filter(langid == "en")

x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
empty <- which(is.na(x$time_until_answer))
for (i in empty) {
  x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}

library(stringr)
library(rebus)
library(survival) 
```

```{r, include = FALSE}
library(qdap)
library(tm)

#function to clean corpus
clean_corpus <- function(corpus){
  library(tm)
  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, content_transformer(tolower))
  corpus <- tm_map(corpus, removeWords, c(stopwords("en"), "can", "will", "cant", "wont", "works", "get", "help", "need"))
  return(corpus)
}

#function that takes in character vector and returns a data frame of words and frequencies sorted by most to least frequent, n is an optional argument 
freq_terms <- function(vec, n = NULL) {
  library(qdap)
  library(tm)
  source <- VectorSource(vec)
  corpus <- VCorpus(source)
  cleaned_corpus <- clean_corpus(corpus) 
  dtm <- DocumentTermMatrix(cleaned_corpus)
  m <- as.matrix(dtm) 
  freq <- colSums(m)
  freq <- sort(freq, decreasing = TRUE)
  freq_df <- data.frame(word = names(freq), frequency = unname(freq))

  if (is.null(n)) {
    return(freq_df)
  }
  else if (!is.null(n)) {
    return(freq_df[1:n,])
  }
}

#function to plot survival curves with ggplot
plot_surv <- function(survfit, data, xlim = NULL) {
  library(ggfortify)
  library(directlabels)

  df <- fortify(survfit, data = data)

  if (missing(xlim)) {

    if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)")
          ggtitle("Survival Curve")
        }
    if ("strata" %in% names(df)) {
      ggplot(df, aes(x = time, y = surv, color = strata)) + 
        geom_line() + 
        scale_y_continuous("Survival Probabilities") +
        scale_x_continuous("Time (hours)") + 
        ggtitle("Survival Curves") + 
        geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
    }
  } else {

      if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)", limits = xlim)
          ggtitle("Survival Curve")
      }

      if ("strata" %in% names(df)) {
        ggplot(df, aes(x = time, y = surv, color = strata)) + 
          geom_line() + 
          scale_y_continuous("Survival Probabilities") +
          scale_x_continuous("Time (hours)", limits = xlim) + 
          ggtitle("Survival Curves") + 
          geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
      }
  }
}
```

### finding top10 and top50 most frequently used terms in titles

```{r, echo = FALSE}
titles_freq <- freq_terms(x$title)

ggplot(titles_freq, aes(x = frequency)) + 
  geom_histogram() + 
  ggtitle("Distribution of word frequencies")
```

```{r, echo = FALSE}
top50 <- titles_freq[1:50,]
top50$word <- factor(top50$word, levels = top50$word)

ggplot(top50, aes(x = word, y = frequency/nrow(x))) + 
  geom_point() +
  scale_y_continuous("Percentage of times the word appears", labels = scales::percent) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ggtitle("How often the 50 most frequently used words appear")
```

##### top10 most frequently used words

```{r, echo = FALSE}
top10 <- titles_freq[1:10,]
x$contain_top10 <- str_detect(as.character(x$title), pattern = or1(names(top10)))
prop_contain <- sum(x$contain_top10)/nrow(x)
```

* `r round(prop_contain * 100, 2)`% of questions contain at least 1 of the top 10 most frequent words in the title

```{r, echo = FALSE}
top10
```

##### median time and average daily views grouped by whether or not question contains top10

```{r, echo = FALSE}
x %>%
  group_by(contain_top10) %>%
  summarise(median_time = median(time_until_answer), 
            avg_daily = mean(daily_views), n = n()) %>%
  arrange(median_time)
```

* table indicates that questions with titles that don't contain the top10 words may get answered faster

##### log-rank test

```{r, echo = FALSE}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
survdiff(surv_object~contain_top10, data = x)
```

```{r, echo = FALSE}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
KM_top10 <- survfit(surv_object ~ contain_top10, data = x)
plot_surv(KM_top10, data = x)
```

### Most frequent words within answered and unanswered questions

```{r, echo = FALSE}
answered <- x %>%
  tbl_df() %>% 
  filter(answered == 1)

unanswered <- x %>%
  tbl_df() %>% 
  filter(answered == 0)

terms_a <- freq_terms(answered$title)
terms_a$prop_in_answered <- terms_a$frequency/nrow(terms_a)
colnames(terms_a)[2] <- "frequency_a"

terms_u <- freq_terms(unanswered$title)
terms_u$prop_in_unanswered <- terms_u$frequency/nrow(terms_u)
colnames(terms_u)[2] <- "frequency_u"

combined <- full_join(terms_a, terms_u, by = "word")
combined$ratio <- combined$prop_in_answered / combined$prop_in_unanswered
head(combined)
```

* data frame with frequency and proportion of times each word appears in both answered and unanswered questions, along with the ratio of proportion of times a word appears in answered questions to the proportion of times it appears in unanswered questions (ratio = prop_in_answered/prop_in_unanswered)

##### viewing proportions/frequencies arranged by ratio in decreasing order

* interpretation: "ssd" is 28.8 times more likely to show up in answered questions than it is in unanswered questions. But since it only shows up 0.7% of the time in answered questions, and 0.02% of the time in unanswered questions, this term might not be helpful in telling us whether or not a question is likely to be answered.

```{r, echo = FALSE}
head(combined %>% 
  arrange(desc(ratio)))
```

##### viewing proportions/frequencies arranged by proportion in unanswered questions, decreasing order 
```{r, echo = FALSE}
head(combined %>%
  arrange(desc(prop_in_unanswered)))
```

```{r, echo = FALSE}
ggplot(combined, aes(x = prop_in_answered)) +
  geom_histogram(bins = 50) + 
  scale_x_continuous("Proportion in answered questions",limits = c(0, 0.05), labels = scales::percent) + 
  scale_y_continuous(limits = c(0,200)) + 
  ggtitle("Distribution of proportions in answered questions")
```

```{r, echo = FALSE}
ggplot(combined, aes(x = prop_in_unanswered)) +
  geom_histogram(bins = 50) + 
  scale_x_continuous("Proportion in unanswered questions", limits = c(0, 0.05), labels = scales::percent) + 
  scale_y_continuous(limits = c(0,200)) + 
  ggtitle("Distribution of proportions in unanswered questions")
```

##### finding proportion and ratio thresholds

* find a set of words that show up frequently enough to be used as a set of predictors
* want the ratio to be large (or small) enough to allow us to decide whether or not that word is more likely to show up in an answered/unanswered question

###### words that appear more than 1% of the time in unanswered questions and are more likely (ratio < 1) to be used in unanswered questions
```{r, echo = FALSE}
p_threshold <- 0.01
ratio_threshold <- 1

freq_terms_u <- combined %>%
                  filter(prop_in_unanswered > p_threshold) %>%
                  filter(ratio < ratio_threshold)
freq_terms_u
```

###### words that appear more than 1% of the time in answered questions and are more likely (ratio > 1) to be used in answered questions

```{r, echo = FALSE}
freq_terms_a <- combined %>%
                  filter(prop_in_answered > p_threshold) %>%
                  filter(ratio > ratio_threshold)
freq_terms_a
```

### Summary statistics/survival analysis for whether or not questions contain frequent unanswered terms

```{r, echo = FALSE}
x$contain_u <- str_detect(as.character(x$title), pattern = or1(freq_terms_u$word))

prop_contain_u <- sum(x$contain_u)/nrow(x)
```

* `r round(prop_contain_u * 100, 2)`% of questions contain at least one of the most frequent unanswered terms

##### median answer time and average daily views grouped by whether or not question contains at least 1 uanswered term

```{r, echo = FALSE}
x %>%
  group_by(contain_u) %>%
  summarise(median_time = median(time_until_answer), n = n(), avg_views = mean(daily_views)) %>%
  arrange(median_time)
```

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ contain_u, data = x)
```

```{r, echo = FALSE}
KM_u <- survfit(surv_object ~ contain_u, data = x) 
plot_surv(KM_u, data = x)
```

### Summary statistics/survival analysis for whether or not questions contain frequent answered terms

```{r, echo = FALSE}
x$contain_a <- str_detect(as.character(x$title), pattern = or1(freq_terms_a$word))

prop_contain_a <- sum(x$contain_a)/nrow(x)
```

* `r round(prop_contain_a * 100, 2)`% of questions contain at least one of the most frequent answered terms

##### median answer time and average daily views grouped by whether or not question contains at least 1 answered term

```{r, echo = FALSE}
x %>%
  group_by(contain_a) %>%
  summarise(median_time = median(time_until_answer), n = n(), avg_views = mean(daily_views)) %>%
  arrange(median_time)
```

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object~contain_a, data = x)
```

```{r, echo = FALSE}
KM_a <- survfit(surv_object ~ contain_a, data = x)
plot_surv(KM_a, data = x)
```

#### looking at only device terms within freq_terms_u (discarding irrelevant words)
```{r, echo = FALSE}
unanswered_devices <- freq_terms_u %>%
                        filter(word %in% c("camera", "sound", "tablet", "light", "wifi", "speaker", "sound"))

x$contain_u_dev <- str_detect(as.character(x$title), pattern = or1(unanswered_devices$word))

prop_contain_u_dev <- sum(x$contain_u_dev)/nrow(x)
```

* `r round(prop_contain_u_dev*100, 2)`% of titles contain at least 1 unanswered device terms

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ contain_u_dev, data = x)

KM_u_dev <- survfit(surv_object ~ contain_u_dev, data = x)
plot_surv(KM_u_dev, data = x)
```

#### Looking at only device terms within freq_terms_a

```{r, echo = FALSE}
answered_devices <- freq_terms_a %>%
                      filter(word %in% c("screen", "iphone", "phone", "battery", "button", "display", "macbook", "lcd", "board", "keyboard", "laptop", "home", "pro", "drive"))

x$contain_dev_a <- str_detect(as.character(x$title), pattern = or1(answered_devices$word))
prop_contain_a_dev <- sum(x$contain_dev_a)/nrow(x)
```

* `r round(prop_contain_a_dev*100, 2)`% of titles contain at least 1 answered device term

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ contain_dev_a, data = x)

KM_dev_a <- survfit(surv_object ~ contain_dev_a, data = x)
plot_surv(KM_dev_a, data = x)
```