### 2017-07-31 tag_exploration.Rmd

---
title: "Exploring n_tags"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, setting up the data, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

library(dplyr)
x <- out %>% 
  tbl_df() %>%
  filter(langid == "en")

x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
empty <- which(is.na(x$time_until_answer))
for (i in empty) {
  x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}

library(stringr)
library(stringi)
library(ggplot2)
library(dplyr)
library(rebus)
library(forcats)
library(purrr)
library(survival)
library(ggfortify)
library(directlabels)
library(htmltools); library(htmlwidgets)
```

```{r, functions, include = FALSE}
plot_surv <- function(survfit, data, xlim) {

  df <- fortify(survfit, data = data)

  if (missing(xlim)) {

    if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)")
          ggtitle("Survival Curve")
        }
    if ("strata" %in% names(df)) {
      ggplot(df, aes(x = time, y = surv, color = strata)) + 
        geom_line() + 
        scale_y_continuous("Survival Probabilities") +
        scale_x_continuous("Time (hours)") + 
        ggtitle("Survival Curves") + 
        geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
    }
  } else {

      if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)", limits = xlim)
          ggtitle("Survival Curve")
      }

      if ("strata" %in% names(df)) {
        ggplot(df, aes(x = time, y = surv, color = strata)) + 
          geom_line() + 
          scale_y_continuous("Survival Probabilities") +
          scale_x_continuous("Time (hours)", limits = xlim) + 
          ggtitle("Survival Curves") + 
          geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
      }
  }
}
```

```{r, echo = FALSE}
#matrix of tags (splitting up tags variable wherever there's a comma)
split_tags <- str_split(x$tags, ", ", simplify = TRUE)

#character vector of all tags (empty strings removed)
tag_vector <- as.vector(split_tags)
tag_vector <- tag_vector[which(tag_vector != "")]

unique_tags <- unique(tag_vector)
num_unique <- length(unique_tags)/length(tag_vector)

prop_notags <- sum(x$n_tags == 0)/nrow(x)
```

* create variable for incorrect tagging 
* create variable for popularity of tags (based off of average freq/popularity of tags)
* `r round(prop_notags*100, 2)`% of questions don't contain any tags
* `r round(num_unique*100, 2)`% of tags are unique

### avg_tag_length

* average length of tag for a question: total number of characters for all of a question's tags / number of tags the question has
* think that tags that are more concise (fewer characters) will have faster answer times

```{r, echo = FALSE}
#average tag length: 
x$avg_tag_length <- NA
not_na <- which(x$tags != "")
for (i in not_na) {
  total_char <- sum(str_length(as.vector(split_tags[i,])))
  total_tags <- sum(as.vector(split_tags[i,]) != "")
  x$avg_tag_length[i] <- total_char / total_tags
}

ggplot(x, aes(x = avg_tag_length)) + 
  geom_histogram() + 
  scale_x_continuous("Average tag length for each question (characters)") + 
  ggtitle("Distribution of average tag lengths")

max_avg <- max(x$avg_tag_length, na.rm = TRUE)
min_avg <- min(x$avg_tag_length, na.rm = TRUE)
mean_length <- mean(x$avg_tag_length, na.rm = TRUE)
median_length <- median(x$avg_tag_length, na.rm = TRUE)
```

* longest average tag length: `r max_avg`
* shortest average tag length: `r min_avg`
* mean average tag length: `r mean_length`
* median average tag length: `r median_length`

##### average tag length grouped by product category 
```{r, echo = FALSE}
x %>%
  group_by(category) %>%
  summarise(avg_length = mean(avg_tag_length, na.rm = TRUE), median_length = median(avg_tag_length, na.rm = TRUE), n = n(), avg_views = mean(daily_views), median_time = median(time_until_answer)) %>%
  arrange(avg_length)
```

##### average tag length grouped by answered/unanswered questions
```{r, echo = FALSE}
x %>%
  group_by(as.factor(answered)) %>%
  summarise(avg_length = mean(avg_tag_length, na.rm = TRUE))
```

### frequency of tags (popularity)

```{r, echo = FALSE}
tag_freq <- data.frame(tag = unique_tags, percent = purrr::map_dbl(unique_tags, ~mean(rowSums(split_tags == .) > 0)))

tag_freq <- tag_freq %>%
              arrange(desc(percent))

ggplot(tag_freq, aes(x = percent)) + 
  geom_histogram() + 
  ggtitle("Distribution of tag frequency")

top50 <- tag_freq[1:50,]
ggplot(top50, aes(x = fct_reorder(tag, -percent), y = percent)) + 
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ggtitle("Distribution of top 50 tags and percentage of times they occur")
```

```{r, echo = FALSE}
#creating average frequency/popular score variable
x$tag1 <- split_tags[,1]
x$tag2 <- split_tags[,2]
x$tag3 <- split_tags[,3]
x$tag4 <- split_tags[,4]

assign_score <- function(data, variable) {
  score <- rep(0, nrow(data))
  notempty <- which(data[[variable]] != "")
  for (i in notempty) {
    score[i] <- tag_freq$percent[which(tag_freq$tag == data[[variable]][i])]
  }
  return(score)
}

x$score1 <- assign_score(x, "tag1")
x$score2 <- assign_score(x, "tag2")
x$score3 <- assign_score(x, "tag3")
x$score4 <- assign_score(x, "tag4")

x$avg_score <- (x$score1 + x$score2 + x$score3 + x$score4)/x$n_tags
x$avg_score[is.nan(x$avg_score)] <- 0

ggplot(x, aes(x = avg_score)) + 
  geom_histogram() + 
  ggtitle("Distribution of average tag popularity/frequency scores")
```

#### Determining threshold for popular/unpopular tags

```{r, echo = FALSE}
median_score <- median(x$avg_score)
q_score <- quantile(x$avg_score, probs = seq(0, 1, by = 0.25))
percentile80 <- quantile(x$avg_score, probs = 0.80)

x$popular_tag <- FALSE
x$popular_tag[x$avg_score >= percentile80] <- TRUE

prop_popular <- sum(x$popular_tag)/sum(x$n_tags != 0)
```

* median average popularity/frequency score: `r median_score` 
* 75th percentile for scores: `r q_score[4]`
* popularity threshold based off of 80th percentile: `r percentile80`
* proportion of questions that contain a tag that are "popular": `r prop_popular`

##### Log-rank test

```{r, echo = FALSE}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
survdiff(surv_object ~ popular_tag, data = x)

KM_popular <- survfit(surv_object ~ popular_tag, data = x)
plot_surv(KM_popular, data = x)
```

### incorrect use of tags
* correct tags
    + don't begin with "#"
    + have no more than 3 words (?)
    + no spelling errors (ex: troubleshoot espressomaschine)
    + no punctuation marks (ex: !, ?, "", /) - since there are tags like: macbook pro 15", just one " is fine
    + should not end with a "."
    + doesn't contain links
    + if it contains numbers must also contain letters 
* regex
* look at probability a quesiton includes certain tag- chances it's answered (grab most frequent)
* maybe ignore tags that are also in device names

```{r, echo = FALSE}
x$correct_tag <- NA_character_
x$correct_tag[x$n_tags != 0] <- TRUE

#function to identify incorrect tagging, single tag taken as input
incorrect_tagging <- function(tag) {
  incorrect <- NA
  if (str_count(tag, pattern = "\\w+") > 3 |
      str_detect(tag, pattern = "[[:punct:]]") | 
      str_detect(tag, pattern = "^[[:digit:]]*$")) {
    incorrect <- TRUE 
  } else {
    incorrect <- FALSE
  }
  return(incorrect)
}

#function to assign t/f for incorrect tags
assign_incorrect <- function(tag_column) {
  new <- rep(NA, nrow(x))
  for (i in which(x[[tag_column]] != "")) {
    new[i] <- incorrect_tagging(x[[tag_column]][i])
  }
  return(new)
}

x$incorrect1 <- assign_incorrect("tag1")
x$incorrect2 <- assign_incorrect("tag2")
x$incorrect3 <- assign_incorrect("tag3")
x$incorrect4 <- assign_incorrect("tag4")

x$correctly_tag <- NA
x$correctly_tag[x$n_tags != 0] <- TRUE
notempty <- which(x$n_tags != 0)
for (k in notempty) {
  if (TRUE %in% c(x$incorrect1[k], x$incorrect2[k], x$incorrect3[k], x$incorrect4[k])) {
    x$correctly_tag[k] <- FALSE
  }
}

prop_correct <- sum(x$correctly_tag, na.rm = TRUE)/sum(x$n_tags != 0)
```

* out of all of the questions that contain tags, `r round(prop_correct * 100, 2)`% of them were tagged correctly

##### Log-rank test

```{r, echo = FALSE}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
survdiff(surv_object ~ correctly_tag, data = x)

KM_correct <- survfit(surv_object ~ correctly_tag, data = x)
plot_surv(KM_correct, data = x)
```

### If a question contains top50 tags

```{r, echo = FALSE}
top50tag <- as.character(top50$tag)

#function to check if tag in top50
contain_tag <- function(tag, data) {
  contain <- rep(NA, nrow(data))
  for(i in which(data[[tag]] != "")) {
    if (data[[tag]][i] %in% top50tag) {
      contain[i] <- TRUE
      } else {
        contain[i] <- FALSE
      }
    }
  return(contain)
}

containdf <- data.frame(c1 = contain_tag("tag1", data = x),
                        c2 = contain_tag("tag2", data = x),
                        c3 = contain_tag("tag3", data = x),
                        c4 = contain_tag("tag4", data = x))

x$contain_toptag <- rep(FALSE, nrow(x))
for (j in which(!is.na(containdf$c1))) {
  if (TRUE %in% containdf[j,]) {
    x$contain_toptag[j] <- TRUE
  } else {
    x$contain_toptag[j] <- FALSE
  }
}

prop_toptag <- sum(x$contain_toptag, na.rm = TRUE)/sum(!is.na(x$tags))
```

* `r round(prop_toptag*100, 2)`% of questions that contain tag, contain a tag that's in the top 50 tag list

##### median time, prop answered and average daily views grouped by whether or not it contains a top50 tag
```{r, echo = FALSE}
x %>% 
  group_by(contain_toptag) %>%
  summarise(median_time = median(time_until_answer), avg_views = mean(daily_views), n = n(), prop_answered = sum(answered)/n)
```

##### log-rank test 
```{r, echo = FALSE}
survdiff(surv_object ~ contain_toptag, data = x)
KM_toptag <- survfit(surv_object ~ contain_toptag, data = x)
plot_surv(KM_toptag, data = x)
```

### maximum number of words in a tag

##### frequency table for number of words

```{r, echo = FALSE}
x$max_tagwords <- rep(NA, nrow(x))
for (i in which(x$n_tags != 0)) {
  tags <- as.vector(split_tags[i,])
  x$max_tagwords[i] <- max(str_count(tags, pattern = "\\w+"))
}

#to find the mode
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

table(x$max_tagwords)
```

* max number of words in a tag: `r max(x$max_tagwords, na.rm = TRUE)`
* median number of words: `r median(x$max_tagwords, na.rm = TRUE)`
* mode number of words: `r Mode(subset(x$max_tagwords, !is.na(x$max_tagwords)))`

##### Average number of max words in a tag grouped by answered/unanswered

```{r, echo = FALSE}
x %>%
  group_by(answered) %>%
  summarise(median_words = median(max_tagwords, na.rm = TRUE))
```