### 2017-07-31 textual_analysis.Rmd

---
title: "Textual Features"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

library(dplyr)
x <- out %>% 
      tbl_df() %>%
      filter(langid == "en")

x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
empty <- which(is.na(x$time_until_answer))
for (i in empty) {
  x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}

library(stringr)
library(rebus)
library(survival) 
library(ggplot2)
library(dplyr)
library(ggfortify)
library(directlabels)
library(openNLP)
```

```{r, include = FALSE}
#function to plot survival curves with ggplot
plot_surv <- function(survfit, data, xlim) {

  df <- fortify(survfit, data = data)

  if (missing(xlim)) {

    if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)")
          ggtitle("Survival Curve")
        }
    if ("strata" %in% names(df)) {
      ggplot(df, aes(x = time, y = surv, color = strata)) + 
        geom_line() + 
        scale_y_continuous("Survival Probabilities") +
        scale_x_continuous("Time (hours)") + 
        ggtitle("Survival Curves") + 
        geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
    }
  } else {

      if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)", limits = xlim)
          ggtitle("Survival Curve")
      }

      if ("strata" %in% names(df)) {
        ggplot(df, aes(x = time, y = surv, color = strata)) + 
          geom_line() + 
          scale_y_continuous("Survival Probabilities") +
          scale_x_continuous("Time (hours)", limits = xlim) + 
          ggtitle("Survival Curves") + 
          geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
      }
  }
}
```

* **question quality:** 
    + text speak: find corpus for text speak- presence/degree of presence/absence of text speak (lol, rofl, u, ur…)

### Length of sentence until first punctuation in text

```{r, echo = FALSE}
x$first_punct <- str_locate(x$text, pattern = "[.|,|?|!|;|:]")[,1]

max_pos <- max(x$first_punct, na.rm = TRUE)
min_pos <- min(x$first_punct, na.rm = TRUE)
mean_pos <- mean(x$first_punct, na.rm = TRUE)
median_pos <- median(x$first_punct, na.rm = TRUE)
prop_punct <- sum(!is.na(x$first_punct))/nrow(x)
```

* `r round(prop_punct*100, 2)`% of questions contain any punctuation marks (. , ? ! ; :)
* max position of first punc: `r max_pos`
* min position of first punc: `r min_pos`
* mean of first_punc: `r mean_pos`
* median of first_punc: `r median_pos`

```{r, echo = FALSE}
ggplot(x, aes(x = first_punct)) + 
  geom_histogram() + 
  ggtitle("Distribution of positions of first punctuation marks") + 
  scale_x_continuous("Position")
```

##### Average and median first_punc grouped by answered

```{r, echo = FALSE}
x %>%
  group_by(answered) %>%
  summarise(avg_pos = mean(first_punct, na.rm = TRUE), median_pos = median(first_punct, na.rm = TRUE), n = n())
```

##### Average and median first_punct grouped by category

* notice that categories that have the shortest distance until first punctuation mark are also the categories who have the fastest time_until_answer values

```{r, echo = FALSE}
x %>%
  group_by(category) %>%
  summarise(avg_pos = mean(first_punct, na.rm = TRUE), median_pos = median(first_punct, na.rm = TRUE), n = n()) %>%
  arrange(median_pos)
```

### Whether or not the text contains punctuation

##### Median time_until_answer grouped by whether or not text contains punct

```{r, echo = FALSE}
x$contain_punct <- FALSE
x$contain_punct[!is.na(x$first_punct)] <- TRUE

x %>%
  group_by(contain_punct) %>%
  summarise(median_time = median(time_until_answer), n = n())
```

##### Proportion of answered/unanswered questions that contain punct

```{r, echo = FALSE}
x %>%
  group_by(answered) %>%
  summarise(n = n(), prop_contain = sum(contain_punct)/n)
```

##### Proportion that contain punct grouped by categories

```{r}
x %>% 
  group_by(category) %>%
  summarise(n = n(), prop_contain = sum(contain_punct)/n, median_pos = median(first_punct, na.rm = TRUE)) %>% arrange(desc(prop_contain))
```

##### Log-rank test

```{r, echo = FALSE}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
survdiff(surv_object ~ contain_punct, data = x)

KM_punct <- survfit(surv_object ~ contain_punct, data = x)
plot_surv(KM_punct, data = x)
```

### Whether or not the text ends in a punctuation mark

```{r, echo = FALSE}
x$end_punct <- str_detect(x$text, pattern = "[.|?|!]$")

prop_end <- sum(x$end_punct)/nrow(x)
```

* `r round(prop_end*100, 2)`% of question texts end with: . ! ?

##### Median time_until_answer grouped by whether or not text ends with punct

```{r, echo = FALSE}
x %>% 
  group_by(end_punct) %>%
  summarise(median_time = median(time_until_answer), n = n())
```

##### Proportion of answered/unanswered questions that end with punct

```{r, echo = FALSE}
x %>% 
  group_by(answered) %>%
  summarise(n = n(), prop_end = sum(end_punct)/n)
```

##### Proportion with end punct grouped by category

```{r, echo = FALSE}
x %>%
  group_by(category) %>%
  summarise(n = n(), prop_end = sum(end_punct)/n) %>%
  arrange(desc(prop_end))
```

##### Log-rank test
```{r, echo = FALSE}
survdiff(surv_object ~ end_punct, data = x)

KM_end <- survfit(surv_object ~ end_punct, data = x)
plot_surv(KM_end, data = x)
```

### Does the title contain a question mark

```{r, question mark, echo = FALSE}
x$contain_question <- str_detect(x$title, pattern = QUESTION)

prop_contain <- sum(x$contain_question)/nrow(x)

true <- x %>% 
  filter(contain_question == T)
prop_answered <- sum(true$answered)/nrow(true)
```
* `r round(prop_contain * 100, 2)`% of questions contain a question mark
* `r round(prop_answered * 100, 2)`% of questions with a question mark were answered

##### median time grouped by whether or not title contains: '?'

```{r, echo = FALSE}
x %>%
  group_by(contain_question) %>%
  summarise(median_time = median(time_until_answer), n = n()) %>%
  arrange(median_time)
```

* indicates that questions with a '?' tend to have shorter answer times than questions without

##### proportion of questions with '?' and median time grouped by category

```{r, echo = FALSE}
x %>% 
  group_by(category) %>%
  summarise(n = n(), proportion_with_question = sum(contain_question)/n, median_time = median(time_until_answer)) %>%
  arrange(desc(proportion_with_question))
```

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ contain_question, data = x)
KM_question <- survfit(surv_object ~ contain_question, data = x)
plot_surv(KM_question, data = x)
```

### If title begins with "Wh"

```{r, echo = FALSE}
x$begin_wh <- str_detect(str_to_lower(x$title), pattern = "^wh")

prop_beginwh <- sum(x$begin_wh)/nrow(x)
```

* `r round(prop_beginwh*100, 2)`% of titles begin with "wh"

##### median time and proportion answered grouped by begin_wh
```{r, echo = FALSE}
x %>%
  group_by(begin_wh) %>%
  summarise(median_time = median(time_until_answer), n = n(), prop_answered = sum(answered)/n)
```

##### log-rank test
```{r, echo = FALSE}
survdiff(surv_object ~ begin_wh, data = x)
KM_wh <- survfit(surv_object ~ begin_wh, data = x)
plot_surv(KM_wh, data = x)
```

### If first letter of title is capitalized

```{r, echo = FALSE}
x$capital_title <- str_detect(x$title, pattern = "^[[:upper:]]")

prop_titlec <- sum(x$capital_title)/nrow(x)
```

* `r round(prop_titlec*100, 2)`% of titles have the first letter capitalized

##### Log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ capital_title, data = x)
KM_titlecaps <- survfit(surv_object ~ capital_title, data = x)
plot_surv(KM_titlecaps, data = x)
```

### If first letter of text is capitalized
```{r, echo = FALSE}
x$capital_text <- str_detect(x$text, pattern = "^[[:upper:]]")

prop_captext <- sum(x$capital_text)/nrow(x)
```

* `r round(prop_captext*100, 2)`% of texts has the first letter capitalized

##### Proportion in answered/unanswered questions with capital letter in text
```{r, echo = FALSE}
x %>%
  group_by(answered) %>%
  summarise(n = n(), prop_capital = sum(capital_text)/n)
```

##### Log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ capital_text, data = x)
KM_textcaps <- survfit(surv_object ~ capital_text, data = x)
plot_surv(KM_textcaps, data = x)
```

### Checking if text is in single caps (all lower or upper case)

* since there are no questions with text in all caps, will only search for all lower case
```{r, echo = FALSE}
x$removed <- str_replace_all(x$text, " ", "")
x$removed <- str_replace_all(x$removed, "[[:punct:]]|[[:digit:]]", "")

x$all_lower <- str_detect(x$removed, pattern = "^[[:lower:]]+$")
x <- x[,-which(names(x) == "removed")]

prop_allLower <- sum(x$all_lower)/nrow(x)
```

* `r round(prop_allLower*100, 2)`% of questions contain text thats all lower case
* this is such a small proportion of questions can we still consider it significant?

##### Proportion all lower case grouped by answered/unanswered

```{r, echo = FALSE}
x %>%
  group_by(answered) %>%
  summarise(n = n(), prop_allLower = sum(all_lower)/n)
```

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ all_lower, data = x)
KM_lower <- survfit(surv_object ~ all_lower, data = x)
plot_surv(KM_lower, data = x)
```

### Searching for words that indicate prior effort in the text variable

```{r, tried, echo = FALSE}
x$prior_effort <- str_detect(str_to_lower(x$text), pattern = or("tried", "searched", "researched", "tested", "replaced", "used", "checked", "investigated", "considered", "measured", "attempted", "inspected", "fitted"))
prop_effort <- sum(x$prior_effort)/nrow(x)
```

* `r round(prop_effort * 100, 2)`% of questions contain terms indicating prior effort

##### median time/average daily views grouped by whether or not question contains "tried"/"searched"
```{r, echo = FALSE}
x %>%
  group_by(prior_effort) %>%
  summarise(median_time = median(time_until_answer), avg_daily = mean(daily_views), n = n()) %>%
  arrange(median_time)
```

##### proportion within answered/unanswered questions that indicate prior effort

```{r, echo = FALSE}
x %>%
  group_by(answered) %>%
  summarise(n = n(), prop_indicateeffort = sum(prior_effort)/n) 
```

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ prior_effort, data = x)
KM_effort <- survfit(surv_object ~ prior_effort, data = x)
plot_surv(KM_effort, data = x)
```

### Do manners matter when asking a question?

```{r, echo = FALSE}
x$gratitude <- str_detect(str_to_lower(x$text), pattern = or("please", "thank you", "thanks", "thankful", "appreciate", "appreciated", "grateful"))

prop_thanks <- sum(x$gratitude)/nrow(x)
```

* `r round(prop_thanks*100,2)`% of questions contain words that express gratitude/manners

##### Median time grouped by whether or not gratitude is expressed

```{r, echo = FALSE}
x %>%
  group_by(gratitude) %>%
  summarise(median_time = median(time_until_answer))
```

##### proportion in answered/unanswered questions that express gratitude

```{r, echo = FALSE}
x %>% 
  group_by(answered) %>%
  summarise(n = n(), prop_gratitude = sum(gratitude)/n)
```

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ gratitude, data = x)
KM_gratitude <- survfit(surv_object ~ gratitude, data = x)
plot_surv(KM_gratitude, data = x)
```

### Does greeting people matter? (presence of "hi")

```{r, echo = FALSE}
x$greeting <- str_detect(str_to_lower(x$text), pattern = START %R% or("hey", "hello", "greetings", "hi"))

prop_greet <- sum(x$greeting)/nrow(x)
```

* `r round(prop_greet*100, 2)`% of questions include greetings

##### log-rank test

```{r, echo = FALSE}
survdiff(surv_object ~ greeting, data = x)
KM_greet <- survfit(surv_object ~ greeting, data = x)
plot_surv(KM_greet, data = x)
```

### Ratio of sentences to question length (number of characters)

* figure out how to deal with "..." and "???" "!!!"
* observation 2247 is funky 
* is it worth it to even consider sentence count?

```{r, echo = FALSE}
x$cleaned <- str_replace_all(as.character(x$text), "\\(.*?\\)", " ")
x$cleaned <- str_replace_all(x$cleaned, "\\[.*?\\]", " ")
x$cleaned <- str_replace_all(x$cleaned, "\\{.*?\\}", " ")
x$cleaned <- str_replace_all(x$cleaned, "\\s?(f|ht)(tp)(s?)(://)([^\\.]*)[\\.|/](\\S*)", " ")
x$cleaned <- str_replace_all(x$cleaned, "([_+a-z0-9-]+(\\.[_+a-z0-9-]+)*@[a-z0-9-]+(\\.[a-z0-9-]+)*(\\.[a-z]{2,14}))", " ")
x$cleaned <- str_replace_all(str_to_lower(x$cleaned), ".com", " ")

x$count <- rep(0, nrow(x))
for (i in 1:nrow(x)) {
  text_split <- as.vector(str_split(x$cleaned[i], pattern = "[.|?|!]", simplify = TRUE))
  text_split <- text_split[which(text_split != "")]
  x$count[i] <- length(text_split)
}

ggplot(x, aes(x = count)) + 
  geom_histogram() + 
  ggtitle("Distribution of number of sentences contained in text") + 
  scale_x_continuous("Sentence Count")

max_count <- max(x$count)
min_count <- min(x$count)
mean_count <- mean(x$count)
median_count <- median(x$count)
prop_one <- sum(x$count == 1)/nrow(x)
```

* max number of sentences: `r max_count`
* minimum number of sentences: `r min_count`
* mean count: `r mean_count`
* median count: `r median_count`
* `r round(prop_one*100, 2)`% of texts have only 1 sentence