### 2017-08-04 calculatingMLEs.Rmd

---
title: "MLEs of Parametric Distributions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

library(dplyr)
x <- out %>% 
  tbl_df() %>%
  filter(langid == "en")

x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
empty <- which(is.na(x$time_until_answer))
for (i in empty) {
  x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}

library(fitdistrplus)
library(flexsurv)
library(survival)
library(mosaic)
library(ggplot2)
```

# Formatting data to work with fitdistrplus package

```{r}
df <- data.frame(left = x$time_until_answer, right = x$time_until_answer)
df[x$answered == 0,2] <- NA

```

# Identifying estimates

```{r}
#these distributions work
fit_logis <- fitdistcens(df, "logis")
summary(fit_logis)

fit_lnorm <- fitdistcens(df, "lnorm")
summary(fit_lnorm)

#this "producedNaNs"
fit_weibull <- fitdistcens(df, "weibull")
summary(fit_weibull)

#outputs Non-positive shape parameter over and over again
fit_llogis <- fitdistcens(df, "llogis", control=list(trace=1, REPORT=1))
summary(fit_llogis)
##output correct?

#exponential doesn't work
fit_exp <- fitdistcens(df, "exp", optim.method = "BFGS", lower = 0, start = NULL)
mledist(df, "exp", optim.method = "BFGS")

```