### 2017-08-05 stackoverflow_comparison.Rmd
set up a CR model similar to one described in literature review, compared results

---
title: "Stack Overflow Comparison"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, importing data, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

library(dplyr)
x <- out %>% 
  tbl_df() %>%
  filter(langid == "en")

x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
empty <- which(is.na(x$time_until_answer))
for (i in empty) {
  x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}

library(stringr)
library(rebus)
library(survival)
```

### Fitting a model similar to the one described in "Min(e)d Your Tags"

##### Variables including:
* avg_score: (describe what the variables are)
* num_pop_tags
* n_images
* text_length
* title_length
* title_questionmark
* title_begin_wh
* is_weekend
* prior_effort

```{r, creating variables, include = FALSE}
#setting up the variables for the model

#tag popularity (frequency)
split_tags <- str_split(x$tags, ", ", simplify = TRUE)
#character vector of all tags (empty strings removed)
tag_vector <- as.vector(split_tags)
tag_vector <- tag_vector[which(tag_vector != "")]
unique_tags <- unique(tag_vector)

tag_freq <- data.frame(tag = unique_tags, percent = purrr::map_dbl(unique_tags, ~mean(rowSums(split_tags == .) > 0)))

tag_freq <- tag_freq %>%
              arrange(desc(percent))

#creating average popularity (frequency) score variable
x$tag1 <- split_tags[,1]
x$tag2 <- split_tags[,2]
x$tag3 <- split_tags[,3]
x$tag4 <- split_tags[,4]

assign_score <- function(data, variable) {
  score <- rep(0, nrow(data))
  notempty <- which(data[[variable]] != "")
  for (i in notempty) {
    score[i] <- tag_freq$percent[which(tag_freq$tag == data[[variable]][i])]
  }
  return(score)
}

x$score1 <- assign_score(x, "tag1")
x$score2 <- assign_score(x, "tag2")
x$score3 <- assign_score(x, "tag3")
x$score4 <- assign_score(x, "tag4")

x$avg_score <- (x$score1 + x$score2 + x$score3 + x$score4)/x$n_tags
x$avg_score[is.nan(x$avg_score)] <- 0

#number of popular tag a question contains
threshold <- 0.005

num_pop <- function(var, threshold) {
  num_pop <- rep(0, nrow(x))
  num_pop[x[[var]] >= threshold] <- 1
  return(num_pop)
}

numpop1 <- num_pop("score1", threshold)
numpop2 <- num_pop("score2", threshold)
numpop3 <- num_pop("score3", threshold)
numpop4 <- num_pop("score4", threshold)

x$num_pop_tags <- numpop1 + numpop2 + numpop3 + numpop4

#text length
x$text_length <- str_length(x$text)

#title length
x$title_length <- str_length(x$title)

#end question mark? 
x$title_questionmark <- str_detect(x$title, pattern = QUESTION %R% END)

#begin with "wh"?
x$title_begin_wh <- str_detect(str_to_lower(x$title), pattern = "^wh")

#is weekend?
x$post_datetime <- as.POSIXct(x$post_date,origin="1970-01-01")
x$post_weekday <- factor(weekdays(x$post_datetime), levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
x$is_weekend <- FALSE
x$is_weekend[x$post_weekday == "Saturday" | x$post_weekday == "Sunday"] <- TRUE

#num_active verb
x$prior_effort <- str_detect(str_to_lower(x$text), pattern = or("tried", "searched", "researched", "tested", "replaced", "used", "checked", "investigated", "considered", "measured", "attempted", "inspected", "fitted"))

x$num_prior_effort <- str_count(str_to_lower(x$text), pattern = or("tried", "searched", "researched", "tested", "replaced", "used", "checked", "investigated", "considered", "measured", "attempted", "inspected", "fitted"))
```

```{r, model, echo = FALSE}
crSO <- coxph(Surv(time_until_answer, answered) ~ avg_score + num_pop_tags + n_images + text_length + title_length + title_questionmark + title_begin_wh + is_weekend + prior_effort, data = x)

summary(crSO)
```

```{r, echo = FALSE}
#function to get medians grouped by variables within the data set (since you can't see it in the boxplot)
#input variable with ""
get_medians <- function(var) {
  if (class(var) != "character" | class(var) != "factor") {
    medians <- x %>%
      group_by(as.factor(x[[var]])) %>%
      summarise(median_time = median(time_until_answer), n = n())
  } else {
    medians <- x %>%
      group_by(x[[var]]) %>%
      summarise(median_time = median(time_until_answer), n = n())    
  }
  return(medians)
}
```

```{r, echo = FALSE}
#boxplots for number of popular/frequent tags a question contains
ggplot(x, aes(x = as.factor(num_pop_tags), y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Number of popular tags in a question") + 
  scale_y_continuous("Answer time (hours)") + 
  ggtitle("Boxplots of answer times grouped by number of popular tags")

get_medians("num_pop_tags")

#logrank test
surv_ob <- Surv(x$time_until_answer, x$answered, type = "right")
survdiff(surv_ob ~ as.factor(num_pop_tags), data = x)
```

```{r, echo = FALSE}
#boxplots for n_images
table(x$n_images)

ggplot(x, aes(x = as.factor(n_images), y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("number of images") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots of answer times grouped by number of images") 
#maybe group 8-13 together

get_medians("n_images")

survdiff(surv_ob ~ as.factor(n_images), data = x)
```

```{r, echo = FALSE}
#creating boxplots for title lengths
#using cut function to divide range of title lengths into intervals 
x$title_ints <- cut(x$title_length, breaks = 5)
table(x$title_ints)

ggplot(x, aes(x = title_ints, y = time_until_answer)) +
  geom_boxplot() + 
  scale_x_discrete("Intervals of title length (characters)") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots of answer times grouped by intervals of title length")

#logrank test
survdiff(surv_ob ~ title_ints, data = x)
```

```{r, echo = FALSE}
#creating boxplots for text lengths
x$text_ints <- cut(x$text_length, breaks = 5)
table(x$text_ints)

ggplot(x, aes(x = text_ints, y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Intervals of text length (characters)") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots of answer times grouped by intervals of text length")

#logrank test
survdiff(surv_ob ~ text_ints, data = x)
```

```{r, echo = FALSE}
#boxplots for prior effort
ggplot(x, aes(x = prior_effort, y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("If the question text indicates prior effort") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots for answer times grouped by whether or not the question indicated prior effort")

get_medians("prior_effort")

ggplot(x, aes(x = as.factor(num_prior_effort), y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Number of prior effort references in a questionss text") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots for answer times grouped by number of prior effort references")

get_medians("num_prior_effort")

#logrank tests
survdiff(surv_ob ~ prior_effort, data = x)
survdiff(surv_ob ~ as.factor(num_prior_effort), data = x)
```

```{r, echo = FALSE}
#boxplots for is_weekend
ggplot(x, aes(x = is_weekend, y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Whether or not the question was posted on the weekend") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots for answer times grouped by weekend or not")

get_medians("is_weekend")

#logrank test
survdiff(surv_ob ~ is_weekend, data = x)
```

```{r, echo = FALSE}
x$wh_q[x$title_questionmark == TRUE & x$title_begin_wh == TRUE] <- 1
x$wh_q[x$title_questionmark == FALSE & x$title_begin_wh == FALSE] <- 2
x$wh_q[x$title_questionmark == TRUE & x$title_begin_wh == FALSE] <- 3
x$wh_q[x$title_questionmark == FALSE & x$title_begin_wh == TRUE] <- 4

#1: title contains both a question mark and begins with "wh"
#2: title doesn't contain a question mark and doesn't begin with "wh"
#3: title contains question mark but doesn't begin with "wh"
#4: title doesn't contain question mark but does start with "wh"

ggplot(x, aes(x = as.factor(wh_q), y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Interaction between title_questionmark and title_begin_wh") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots for answer times grouped by levels of interaction")

ggplot(x, aes(x = title_questionmark, y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Whether or not the title contains a question mark") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots for answer times grouped by question mark")

ggplot(x, aes(x = title_begin_wh, y = time_until_answer)) + 
  geom_boxplot() + 
  scale_x_discrete("Whether or not the title begins with 'Wh'") + 
  scale_y_continuous("Time until answer (hours)") + 
  ggtitle("Boxplots for answer times grouped by 'Wh'")

get_medians("wh_q")
get_medians("title_questionmark")
get_medians("title_begin_wh")

#logrank tests
survdiff(surv_ob ~ as.factor(wh_q), data = x)
survdiff(surv_ob ~ title_questionmark, data = x)
survdiff(surv_ob ~ title_begin_wh, data = x)
```

# STILL NEED TO DO 
* figure out what's going on with the HR for avg_score
* compare how these results are different from stack overflows, why are they different?
* make boxplots easier to read, find a way to actually see the bottom halves of boxplots, ignore outliers? 
* for boxplots that only have 1 observation in them, figure out how to group it with a diff value? 
* make this report easier to read