### 2017-08-08 functions.Rmd
reorganized files

---
title: "Functions Created"
author: "Lisa Oshita"
date: "July 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tm)
library(qdap)
```

```{r}
#function to import and set up the data 
setup <- function() {
  dir <- file.path(getwd(),"data")
  out <- read.csv(file.path(dir, "answers_data.csv"))

  library(dplyr)
  x <- out %>% 
    tbl_df() %>%
    filter(langid == "en")

  x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
  empty <- which(is.na(x$time_until_answer))
  for (i in empty) {
    x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}
}
```

```{r}
#functions used in exploratory analysis

#function to plot survfit objects
#input: survfit object, data, optional x-axis limits
plot_surv <- function(survfit, data, xlim = NULL) {
  library(ggfortify)
  library(directlabels)

  df <- fortify(survfit, data = data)

  if (missing(xlim)) {

    if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)")
          ggtitle("Survival Curve")
        }
    if ("strata" %in% names(df)) {
      ggplot(df, aes(x = time, y = surv, color = strata)) + 
        geom_line() + 
        scale_y_continuous("Survival Probabilities") +
        scale_x_continuous("Time (hours)") + 
        ggtitle("Survival Curves") + 
        geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
    }
  } else {

      if (!("strata" %in% names(df))) {
        ggplot(df, aes(x = time, y = surv)) + 
          geom_line() + 
          scale_y_continuous("Survival Probability") +
          scale_x_continuous("Time (hours)", limits = xlim)
          ggtitle("Survival Curve")
      }

      if ("strata" %in% names(df)) {
        ggplot(df, aes(x = time, y = surv, color = strata)) + 
          geom_line() + 
          scale_y_continuous("Survival Probabilities") +
          scale_x_continuous("Time (hours)", limits = xlim) + 
          ggtitle("Survival Curves") + 
          geom_dl(aes(label = strata), method = list(dl.trans(x = x + 0.2), "last.points", cex = 0.5))
      }
  }
}

#function that prints out summary stats grouped by that variable 
sum_stats <- function(variable, data) {
  tbl <- x %>%
    group_by(data[[variable]]) %>%
    summarise(n = n(), median_time = median(time_until_answer), median_views = median(daily_views), prop_answered = sum(answered)/n) %>%
    arrange(median_time)
  names(tbl)[1] <- variable
  return(tbl)
}
```

```{r}
#function to clean corpus, can specify stopwords to remove 
clean_corpus <- function(corpus, stopwrds = NULL){
  if (is.null(stopwrds)) {
    corpus <- tm_map(corpus, stripWhitespace)
    corpus <- tm_map(corpus, content_transformer(tolower))
    corpus <- tm_map(corpus, removeWords, c(stopwords("en"), "can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
    return(corpus)    
  } else {
    corpus <- tm_map(corpus, stripWhitespace)
    corpus <- tm_map(corpus, content_transformer(tolower))
    corpus <- tm_map(corpus, removeWords, c(stopwords("en"), "can", "will", "cant", "wont", "works", "get", "help", "need", "fix", stopwrds))
    return(corpus)
  }
}

#function that takes in character vector and returns a data frame of words and frequencies sorted by most to least frequent, n and stopwrds are an optional arguments
freq_terms <- function(vec, n = NULL, stopwrds = NULL) {
  source <- VectorSource(vec)
  corpus <- VCorpus(source)

  if (is.null(stopwrds)) {
    cleaned_corpus <- clean_corpus(corpus) 
    dtm <- DocumentTermMatrix(cleaned_corpus, control = list(weighting = weightTfIdf))
    m <- as.matrix(dtm) 
    freq <- colSums(m)
    freq <- sort(freq, decreasing = TRUE)
    freq_df <- data.frame(word = names(freq), frequency = unname(freq))    
  } else {
    cleaned_corpus <- clean_corpus(corpus, stopwrds) 
    dtm <- DocumentTermMatrix(cleaned_corpus, control = list(weighting = weightTfIdf))
    m <- as.matrix(dtm) 
    freq <- colSums(m)
    freq <- sort(freq, decreasing = TRUE)
    freq_df <- data.frame(word = names(freq), frequency = unname(freq))
  }

  if (missing(n)) {
    return(freq_df)
  }
  else {
    return(freq_df[1:n,])
  }
}
```

```{r}
#functions used in model fitting

#function to calculate r-square adjusted
#k = number of independent predictors 
get_rsq_adj <- function(cr_obj, k) {
  r <- unname(summary(cr_obj)$rsq[1])
  n <- summary(cr_obj)$n
  radj <- 1- (((1- (r**2)) * (n - 1))/(n-k-1))
  return(radj)
}

#add in function to compare nested models

```

```{r}
#functions used to compare stackoverflow's model

#function to get median answer times grouped by specified variables (since you can't see it in the boxplot)
#input var using ""
get_medians <- function(var) {
  if (class(var) != "character" | class(var) != "factor") {
    medians <- x %>%
      group_by(as.factor(x[[var]])) %>%
      summarise(median_time = round(median(time_until_answer), 2), n = n())
  } else {
    medians <- x %>%
      group_by(x[[var]]) %>%
      summarise(median_time = round(median(time_until_answer), 2), n = n())    
  }
  names(medians)[1] <- var
  return(medians)
}

#function to create boxplot 
get_boxplot <- function(var, xaxis, title) {
  if (class(var) != "character" | class(var) != "factor") {
    ggplot(x, aes(x = as.factor(x[[var]]), y = time_until_answer)) + 
      geom_boxplot() + 
      scale_x_discrete(xaxis) + 
      scale_y_continuous("Answer time (hours)") + 
      ggtitle(paste("Boxplots of answer times grouped by", title))
  } else {
    ggplot(x, aes(x = x[[var]], y = time_until_answer)) + 
      geom_boxplot() + 
      scale_x_discrete(xaxis) + 
      scale_y_continuous("Answer time (hours)") + 
      ggtitle(paste("Boxplots of answer times grouped by", title))
  }
}
```