### 2017-08-15 model fitting/univariate_analysis.Rmd
created new markdown for univariate analysis, stratified model on new_category

---
title: "Univariate Analysis"
author: "Lisa Oshita"
date: "8/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(ggplot2)
```

### Setting up cross validation paths

* cross validation paths stored in splitPlan
* set seed for reproducible results

```{r, echo = FALSE}
set.seed(444)
splitPlan <- vtreat::kWayCrossValidation(nrow(x), 5, NULL, NULL)
str(splitPlan)
```

### Analysis on test data

```{r, data and text variables, include = FALSE}
train1 <- x[splitPlan[[1]]$train, ]

pvalues <- data.frame(variable = character(), pvalue = numeric())

#function to get summary of cox model + append pvalue of variable from loglik test to data frame
get_pvalue <- function(variable, factor = FALSE) {
  if (factor == FALSE){
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ train1[[variable]], data = train1)
  } else {
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ as.factor(train1[[variable]]), data = train1)
  }
  summary <- summary(coxmodel)
  print(summary)
  df <- data.frame(variable = variable, pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}

#===================================================================
# Univariate analyses (select variables with p-value < 0.01)
# category 
train1$category <- as.character(train1$category)
train1$category[is.na(train1$category)] <- "Other"
train1$category <- as.factor(train1$category)
pvalues <- get_pvalue("category")
#===================================================================
# n_images
pvalues <- get_pvalue("n_images")

# new_n_images (recoding to group questions with >= 6 images) 
train1$new_n_images <- as.character(train1$n_images)
train1$new_n_images[train1$n_images >= 6] <- ">= 6"
pvalues <- get_pvalue("new_n_images")

#===================================================================
# n_tags
pvalues <- get_pvalue("n_tags")

#===================================================================
# new_user
pvalues <- get_pvalue("new_user", factor = TRUE)

#===================================================================
# weekday
train1$datetime <- as.POSIXct(train1$post_date,origin="1970-01-01")
train1$weekday <- factor(weekdays(train1$datetime), levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
pvalues <- get_pvalue("weekday")

#===================================================================
# is_weekend
train1$is_weekend <- FALSE
train1$is_weekend[train1$weekday == "Saturday" | train1$weekday == "Sunday"] <- TRUE
pvalues <- get_pvalue("is_weekend")

#===================================================================
# hour of the day
train1$hour <- as.numeric(format(train1$datetime,"%H"))
pvalues <- get_pvalue("hour", factor = TRUE)

#===================================================================
# morning/noon/evening/night?
train1$ampm <- "Night"
train1$ampm[train1$hour >= 5 & train1$hour < 12] <- "Morning"
train1$ampm[train1$hour >= 12 & train1$hour < 17] <- "Afternoon" #noon - 5pm
train1$ampm[train1$hour >= 17 & train1$hour < 20] <- "Evening" #5pm - 8pm
pvalues <- get_pvalue("ampm")

#===================================================================
# title length
train1$title_length <- str_length(train1$title)
pvalues <- get_pvalue("title_length", factor = FALSE)

#===================================================================
# text length
train1$text_length <- str_length(train1$text)
pvalues <- get_pvalue("text_length")

#===================================================================
# device name length
train1$device_length <- str_length(train1$device)
pvalues <- get_pvalue("device_length")

#===================================================================
# length of sentence until first punctuation mark
train1$length <- str_locate(train1$text, pattern = "[.|?|!]")[,1]
q <- stats::quantile(train1$length, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
train1$text_till_punct <- "none"
train1$text_till_punct[train1$length <= q[3]] <- "short"
train1$text_till_punct[train1$length >= q[3] & train1$length <= q[4]] <- "medium"
train1$text_till_punct[train1$length >= q[4]] <- "long"
pvalues <- get_pvalue("text_till_punct")

#===================================================================
# if text contains any end punctuation
train1$text_contain_punct <- str_detect(train1$text, pattern = "[.|?|!]")
pvalues <- get_pvalue("text_contain_punct")

#===================================================================
# if the text ends in a punctuation
train1$text_end_punct <- str_detect(train1$text, pattern = "[.|?|!]$")
pvalues <- get_pvalue("text_end_punct")

#===================================================================
# if the title ends with a question mark
train1$title_questionmark <- str_detect(train1$title, pattern = QUESTION %R% END)
pvalues <- get_pvalue("title_questionmark")

#===================================================================
# if title begins with "wh"
train1$title_beginwh <- str_detect(str_to_lower(train1$title), pattern = "^wh")
pvalues <- get_pvalue("title_beginwh")

#===================================================================
# if first letter of title is capitalized
train1$capital_title <- str_detect(as.character(train1$title), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_title")

#===================================================================
# if first letter of text is capitalized
train1$capital_text <- str_detect(as.character(train1$text), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_text")

#===================================================================
# if text is in all lower case
train1$cleaned <- str_replace_all(train1$text, " ", "")
train1$cleaned <- str_replace_all(train1$cleaned, "[[:punct:]]|[[:digit:]]", "")
train1$text_all_lower <- str_detect(train1$cleaned, pattern = "^[[:lower:]]+$")
pvalues <- get_pvalue("text_all_lower")

#===================================================================
# prior_effort
prior_terms <- c("tried", "searched", "researched", "tested", "replaced", "used", "checked", "investigated", "considered", "measured", "attempted", "inspected", "fitted")
train1$prior_effort <- str_detect(str_to_lower(train1$text), pattern = or1(prior_terms))
pvalues <- get_pvalue("prior_effort")

#===================================================================
# num_prior_effort
train1$num_prior_effort <- str_count(str_to_lower(train1$text), pattern = or1(prior_terms))
pvalues <- get_pvalue("num_prior_effort")

#===================================================================
# expressing gratitude 
gratitude_terms <- c("please", "thank you", "thanks", "thankful", "appreciate", "appreciated", "grateful")
train1$gratitude <- str_detect(str_to_lower(train1$text), pattern = or1(gratitude_terms))
pvalues <- get_pvalue("gratitude")

#===================================================================
# greeting
greeting_terms <- c("hey", "hello", "greetings", "hi")
train1$greeting <- str_detect(str_to_lower(train1$text), pattern = START %R% or1(greeting_terms))
pvalues <- get_pvalue("greeting")

#===================================================================
# update the question?
train1$update <- str_detect(train1$text, pattern = "===")
pvalues <- get_pvalue("update")

#===================================================================
# newline ratio to length of text
train1$newline_ratio <- str_count(train1$text, pattern = "\n")/str_length(train1$text)
pvalues <- get_pvalue("newline_ratio")
```

```{r, tag based and frequent term variables, include = FALSE}
# tag-based variables
#===================================================================
# avg_tag_length
split_tags <- str_split(train1$tags, ", ", simplify = TRUE)
train1$avg_tag_length <- NA
not_na <- which(train1$tags != "")
for (i in not_na) {
  total_char <- sum(str_length(as.vector(split_tags[i,])))
  total_tags <- sum(as.vector(split_tags[i,]) != "")
  train1$avg_tag_length[i] <- total_char / total_tags
}
pvalues <- get_pvalue("avg_tag_length")

#===================================================================
# maximum number of words for a question's tags
train1$max_tagwords <- rep(0, nrow(train1))
for (i in which(train1$n_tags != 0)) {
  tags <- as.vector(split_tags[i,])
  train1$max_tagwords[i] <- max(str_count(tags, pattern = "\\w+"))
}
pvalues <- get_pvalue("max_tagwords")

#===================================================================
# frequent tags
tag_vector <- as.vector(split_tags)
tag_vector <- tag_vector[which(tag_vector != "")]
unique_tags <- unique(tag_vector)

tag_freq <- data.frame(tag = unique_tags, percent = purrr::map_dbl(unique_tags, ~mean(rowSums(split_tags == .) > 0)))
tag_freq <- tag_freq %>%
              arrange(desc(percent))

#creating average frequency score variable
train1$tag1 <- split_tags[,1]
train1$tag2 <- split_tags[,2]
train1$tag3 <- split_tags[,3]
train1$tag4 <- split_tags[,4]

assign_score <- function(data, variable) {
  score <- rep(0, nrow(data))
  notempty <- which(data[[variable]] != "")
  for (i in notempty) {
    score[i] <- tag_freq$percent[which(tag_freq$tag == data[[variable]][i])]
  }
  return(score)
}
train1$score1 <- assign_score(train1, "tag1")
train1$score2 <- assign_score(train1, "tag2")
train1$score3 <- assign_score(train1, "tag3")
train1$score4 <- assign_score(train1, "tag4")
train1$avg_tag_score <- (train1$score1 + train1$score2 + train1$score3 + train1$score4)/as.numeric(train1$n_tags)
train1$avg_tag_score[is.nan(train1$avg_tag_score)] <- 0

pvalues <- get_pvalue("avg_tag_score")

#===================================================================
# if question contains "frequent" tag
percentile80 <- quantile(train1$avg_tag_score, probs = 0.80)

train1$frequent_tag <- FALSE
train1$frequent_tag[train1$avg_tag_score >= percentile80] <- TRUE

pvalues <- get_pvalue("frequent_tag")

#===================================================================
# number of "frequent" tags a question contains
threshold <- 0.005
num_pop <- function(var, threshold) {
  num_pop <- rep(0, nrow(train1))
  num_pop[train1[[var]] >= threshold] <- 1
  return(num_pop)
}
numpop1 <- num_pop("score1", threshold)
numpop2 <- num_pop("score2", threshold)
numpop3 <- num_pop("score3", threshold)
numpop4 <- num_pop("score4", threshold)
train1$num_freq_tags <- numpop1 + numpop2 + numpop3 + numpop4

pvalues <- get_pvalue("num_freq_tags")

#==================================================================================
#frequent terms in unanswered/answered questions
answered <- train1 %>%
  tibble::as_tibble() %>% 
  filter(answered == 1)

unanswered <- train1 %>%
  tibble::as_tibble() %>% 
  filter(answered == 0)

terms_a <- oshitar::get_freq_terms(answered$title, stopwords = c("can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
terms_a$prop_in_answered <- terms_a$frequency/nrow(terms_a)
colnames(terms_a)[2] <- "frequency_a"

terms_u <- oshitar::get_freq_terms(unanswered$title, stopwords = c("can", "will", "cant", "wont", "works", "get", "help", "need", "fix"))
terms_u$prop_in_unanswered <- terms_u$frequency/nrow(terms_u)
colnames(terms_u)[2] <- "frequency_u"

combined <- dplyr::full_join(terms_a, terms_u, by = "word")
combined$ratio <- combined$prop_in_answered / combined$prop_in_unanswered

p_threshold <- 0.01
ratio_threshold <- 1

freq_terms_u <- combined %>%
                  filter(prop_in_unanswered > p_threshold) %>%
                  filter(ratio < ratio_threshold)
freq_terms_a <- combined %>%
                  filter(prop_in_answered > p_threshold) %>%
                  filter(ratio > ratio_threshold)

train1$contain_unanswered <- str_detect(as.character(train1$title), pattern = or1(freq_terms_u$word))
train1$contain_answered <- str_detect(as.character(train1$title), pattern = or1(freq_terms_a$word))

pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")
```

### examining univariate analyses results

```{r, echo = FALSE}
options(scipen=999)
(pvalues <- pvalues %>%
  arrange(pvalue))
```