### 2017-08-17 literature reviews/literature_reviews.tex
updated lit file

\documentclass[12pt]{article}
\usepackage{natbib}  % used for citations
\usepackage[parfill]{parskip} %used for formatting style of text

% Sets margins to 1 in
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%-----------------------------------------------------------------------------------
\title{Literature Review of Online Question Forums}
\author{Lisa Oshita}
\date{}
%-----------------------------------------------------------------------------------

\begin{document}

\maketitle

\section{Introduction}

This really cool paper about stackoverflow said something cool \citep{Ponzanelli2014}.

\cite{Ponzanelli2014} said this awesome thing.

\bibliographystyle{ECA_jasa}
\bibliography{questions}

\end{document}