### 2017-08-21 literature reviews/literature_reviews.tex
set up/learned about tex and mendeley

\documentclass[12pt]{article}
\usepackage{natbib}  % used for citations
\usepackage[parfill]{parskip} %used for formatting style of text

% Sets margins to 1 in
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%-----------------------------------------------------------------------------------
\title{Literature Review of Online Question Forums}
\author{Lisa Oshita}
\date{}
%-----------------------------------------------------------------------------------

\begin{document}

\maketitle

\section{Introduction}

This really cool paper about stackoverflow said something cool \citep{Ponzanelli2014a}.

\cite{Ponzanelli2014a} said this awesome thing.

chen said blah  \citep{Chen}

\bibliographystyle{ECA_jasa}
\bibliography{questions}

\end{document}