### 2017-08-22 model fitting/univariate_analysis.Rmd
univariate analysis for quadratic/cubic terms, square root transformations, splines

---
title: "Univariate Analysis"
author: "Lisa Oshita"
date: "8/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(ggplot2)
```

### Setting up cross validation paths

* cross validation paths stored in splitPlan
* set seed for reproducible results

```{r, echo = FALSE}
set.seed(444)
splitPlan <- vtreat::kWayCrossValidation(nrow(x), 5, NULL, NULL)
str(splitPlan)
```

### Analysis on test data

```{r, data and text variables, include = FALSE}
train1 <- x[splitPlan[[1]]$train, ]

pvalues <- data.frame(variable = character(), pvalue = numeric())

#function to get summary of cox model + append pvalue of variable from loglik test to data frame
get_pvalue <- function(variable, factor = FALSE) {
  if (factor == FALSE){
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ train1[[variable]], data = train1)
  } else {
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ as.factor(train1[[variable]]), data = train1)
  }
  summary <- summary(coxmodel)
  df <- data.frame(variable = variable, pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}

# Univariate analyses (select variables with p-value < 0.01)
#===================================================================
# category 
pvalues <- get_pvalue("category")
#===================================================================
# new_category
pvalues <- get_pvalue("new_category")
#===================================================================
# n_images
pvalues <- get_pvalue("n_images")

# new_n_images (recoding to group questions with >= 6 images) 
train1$new_n_images <- as.character(train1$n_images)
train1$new_n_images[train1$n_images >= 6] <- ">= 6"
pvalues <- get_pvalue("new_n_images")

#===================================================================
# n_tags
pvalues <- get_pvalue("n_tags", factor = TRUE)

#===================================================================
# new_user
pvalues <- get_pvalue("new_user")

#===================================================================
# weekday
pvalues <- get_pvalue("weekday")

#===================================================================
# hour of the day
pvalues <- get_pvalue("hour", factor = TRUE)

#===================================================================
# morning/noon/evening/night?
pvalues <- get_pvalue("ampm")

#===================================================================
# title length
pvalues <- get_pvalue("title_length")

#===================================================================
# text length
pvalues <- get_pvalue("text_length")

#===================================================================
# device name length
pvalues <- get_pvalue("device_length")

#===================================================================
# length of sentence until first punctuation mark
pvalues <- get_pvalue("text_till_punct")

#===================================================================
# if text contains any end punctuation
pvalues <- get_pvalue("text_contain_punct")

#===================================================================
# if the title ends with a question mark
pvalues <- get_pvalue("title_questionmark")

#===================================================================
# if title begins with "wh"
pvalues <- get_pvalue("title_beginwh")

#===================================================================
# if first letter of title is capitalized
train1$capital_title <- str_detect(as.character(train1$title), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_title")

#===================================================================
# if first letter of text is capitalized
train1$capital_text <- str_detect(as.character(train1$text), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_text")

#===================================================================
# if text is in all lower case
pvalues <- get_pvalue("text_all_lower")

#===================================================================
# prior_effort
pvalues <- get_pvalue("prior_effort")

#===================================================================
# expressing gratitude 
pvalues <- get_pvalue("gratitude")

#===================================================================
# greeting
pvalues <- get_pvalue("greeting")

#===================================================================
# update the question?
pvalues <- get_pvalue("update")

#===================================================================
# newline ratio to length of text
pvalues <- get_pvalue("newline_ratio")
```

```{r, tag based and frequent term variables, include = FALSE}
# tag-based variables 
#===================================================================
# avg_tag_length
pvalues <- get_pvalue("avg_tag_length")

#===================================================================
# avg_tag_score
pvalues <- get_pvalue("avg_tag_score")

#===================================================================
# number of "frequent" tags a question contains
pvalues <- get_pvalue("num_freq_tags")

#==================================================================================
#frequent terms in unanswered/answered questions
pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")
```

### researching splines + quadratic terms + square root transformations

* spline: piecewise m - 1 degree polynomial that is continuous up to its first m - 2 derivatives
* obtain more flexible curves: increase the degree of spline/add more knots
* tradeoff: 
    + few knots/low degree: class of functions may be too restrictive (bias)
    + many knots/high degree: risk overfitting 
* spline has m + K degrees of freedom 
* types of splines
    + cubic splines (most common)
* truncated power basis - flawed
* B-spline basis: much more stable, typically used 
* Natural cubic splines
    + polynomial fits tend to be erratic at boundaries of data (even worse for cubic splines)
    + natural cubic splines fix this by adding 4 constraints that the function is linear beyond the boundaries of the data 
    + because basis functions are fixed, model fitting and inference are straightforward exentsions to how the model would normally be without splines 
* **use pspline and penalized spline smoothing method with AIC criteria**

```{r, echo = FALSE}
# fit splines + sqrt/log transformations on title_length, text_length, device_length, avg_tag_length, newline_ratio

# new function to add p-values from these fits to pvalue data frame
get_pvalue_new <- function(variable, fit) {
  summary <- summary(fit)
  df <- data.frame(variable = variable, pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}

#===================================================================

# title_length
spline_title <- coxph(Surv(time_until_answer, answered) ~ pspline(title_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("title_length_spline", fit = spline_title)

sqrt_title <- coxph(Surv(time_until_answer, answered) ~ sqrt(title_length), data = train1)
pvalues <- get_pvalue_new("title_length_sqrt", fit = sqrt_title)

# quadratic/cubic terms

title2 <- coxph(Surv(time_until_answer, answered) ~ title_length + I(title_length^2), data = train1)
pvalues <- get_pvalue_new("quad_title", fit = title2)

title3 <- coxph(Surv(time_until_answer, answered) ~ title_length + I(title_length^2) + I(title_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_title", fit = title3)

#===================================================================

# text_length
spline_text <- coxph(Surv(time_until_answer, answered) ~ pspline(text_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("text_length_spline", fit = spline_text)

sqrt_text <- coxph(Surv(time_until_answer, answered) ~ sqrt(text_length), data = train1)
pvalues <- get_pvalue_new("text_length_sqrt", fit = sqrt_text)

logtext <- coxph(Surv(time_until_answer, answered) ~ log10(text_length), data = train1)
pvalues <- get_pvalue_new("text_length_log10", fit = logtext)

# quadratic/cubic terms

text2 <- coxph(Surv(time_until_answer, answered) ~ text_length + I(text_length^2), data = train1)
pvalues <- get_pvalue_new("quad_text", fit = text2)

text3 <- coxph(Surv(time_until_answer, answered) ~ text_length + I(text_length^2) + I(text_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_text", fit = text3)

#===================================================================

# device_length
spline_dev <- coxph(Surv(time_until_answer, answered) ~ pspline(device_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("device_length_spline", fit = spline_dev)

sqrt_dev <- coxph(Surv(time_until_answer, answered) ~ sqrt(device_length), data = train1)
pvalues <- get_pvalue_new("device_length_sqrt", fit = sqrt_dev)

# quadratic/cubic terms

dev2 <- coxph(Surv(time_until_answer, answered) ~ device_length + I(device_length^2), data = train1)
pvalues <- get_pvalue_new("quad_dev", fit = dev2)

dev3 <- coxph(Surv(time_until_answer, answered) ~ device_length + I(device_length^2) + I(device_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_dev", fit = dev3)

#===================================================================

# newline_ratio
fit7 <- coxph(Surv(time_until_answer, answered) ~ pspline(newline_ratio, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("newline_ratio_spline", fit = fit7)

fit8 <- coxph(Surv(time_until_answer, answered) ~ sqrt(newline_ratio), data = train1)
pvalues <- get_pvalue_new("newline_ratio_sqrt", fit = fit8)

# quadratic/cubic terms

newline2 <- coxph(Surv(time_until_answer, answered) ~ newline_ratio + I(newline_ratio^2), data = train1)
pvalues <- get_pvalue_new("quad_newline", fit = newline2)

newline3 <- coxph(Surv(time_until_answer, answered) ~ newline_ratio + I(newline_ratio^2) + I(newline_ratio^3), data = train1)
pvalues <- get_pvalue_new("cubic_newline", fit = newline3)

#===================================================================

# avg_tag_length
fit9 <- coxph(Surv(time_until_answer, answered) ~ pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("avg_tag_length_spline", fit = fit9)

fit10 <- coxph(Surv(time_until_answer, answered) ~ sqrt(avg_tag_length), data = train1)
pvalues <- get_pvalue_new("avg_tag_length_sqrt", fit = fit10)

# quadratic/cubic terms

tag2 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_length + I(avg_tag_length^2), data = train1)
pvalues <- get_pvalue_new("quad_tag", fit = tag2)

tag3 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_length + I(avg_tag_length^2) + I(avg_tag_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_tag", fit = tag3)

#===================================================================

# avg_tag_score
fit11 <- coxph(Surv(time_until_answer, answered) ~ pspline(avg_tag_score, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("avg_tag_score_spline", fit = fit11)

fit12 <- coxph(Surv(time_until_answer, answered) ~ sqrt(avg_tag_score), data = train1)
pvalues <- get_pvalue_new("avg_tag_score_sqrt", fit = fit12)

# quadratic/cubic terms

score2 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_score + I(avg_tag_score^2), data = train1)
pvalues <- get_pvalue_new("quad_score", fit = score2)

score3 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_score + I(avg_tag_score^2) + I(avg_tag_score^3), data = train1)
pvalues <- get_pvalue_new("cubic_score", fit = score3)
```

### examining univariate analyses results

```{r, echo = FALSE}
options(scipen=999)
(pvalues <- pvalues %>%
  arrange(pvalue))
```