### 2017-08-26 model fitting/cross_validation.Rmd
model fitting

---
title: "Predictive modeling with cross validation (this file is a mess)"
author: "Lisa Oshita"
date: "8/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
x <- oshitar::setup()
library(survival); library(stringr); library(rebus); library(dplyr); library(stats); library(rms)
```

```{r, cross validation, echo = FALSE}
# creating cross-validation plan with k = 5
set.seed(444)
splitPlan <- vtreat::kWayCrossValidation(nrow(x), 5, NULL, NULL)

train1 <- x[splitPlan[[1]]$train, ]; test1 <- x[splitPlan[[1]]$app, ]

train2 <- x[splitPlan[[2]]$train, ]; test2 <- x[splitPlan[[2]]$app, ]

train3 <- x[splitPlan[[3]]$train, ]; test3 <- x[splitPlan[[3]]$app, ]

train4 <- x[splitPlan[[4]]$train, ]; test4 <- x[splitPlan[[4]]$app, ]

train5 <- x[splitPlan[[5]]$train, ]; test5 <- x[splitPlan[[5]]$app, ]
```

### Building the model on one of the training data sets

##### Fitting model with all significant variables

* univariate analysis p-value < 0.01
* using stepAIC to perform backwards and forwards stepwise variable selection with AIC criteria

```{r, echo = FALSE}
# fitted with penalized spline smoothing with AIC criteria on device_length and avg_tag_length
fit <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + log10(text_length) + title_questionmark + title_beginwh + sqrt(newline_ratio) + sqrt(avg_tag_score) + text_contain_punct + text_all_lower + update + greeting + as.factor(n_tags) + gratitude + prior_effort + ampm + n_images + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)

#model <- MASS::stepAIC(fit, direction = "both", k = 2)

# results of stepAIC (including contain_answered)
results <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + 
    contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + 
    text_all_lower + update + prior_effort + ampm + weekday + 
    pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, 
    df = 0, method = "AIC"), data = train1)
AIC(results2, k = 2) # 64866.93
```

##### Splines

* adding in splines one at a time

```{r}
# spline on device_length
spline <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC"), data = train1)
AIC(spline, k = 2) # 64907.21
summary(spline)

# spline on text_length
spline1 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(text_length, df = 0, method = "AIC"), data = train1)
AIC(spline1, k = 2) # 64914.27

# spline on avg_tag_length
spline2 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
AIC(spline2, k = 2) # 64893.52

# spline on avg_tag_length + device_length
spline3 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
AIC(spline3, k = 2) # 64865.95

# spline on newline_ratio
spline4 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(newline_ratio, df = 0, method = "AIC"), data = train1)
AIC(spline4, k = 2) # 64915.6

# log10(text_length), spline = avg_tag_length + device_length
spline5 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
AIC(spline5, k = 2) # 64865.73

# log10(text_length), sqrt(newline_ratio), spline = avg_tag_length + device_length
spline6 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
AIC(spline6, k = 2) # 64864.73

# title_beginwh*title_questionmark, log10(text_length), sqrt(newline_ratio), spline = avg_tag_length + device_length
spline7 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC") + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(spline7, k = 2) # 64858.68

# title_beginwh*title_questionmark, log10(text_length), sqrt(newline_ratio), spline = avg_tag_length + device_length + newline_ratio
spline8 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC") + title_beginwh + pspline(newline_ratio, df = 0, method = "AIC") + title_questionmark*title_beginwh, data = train1)
AIC(spline8, k = 2) # 64841.25
```

##### Restricted cubic splines

* using rcs (restricted cubic spline) in rms package
* fit knots at quantiles of the predictor 
* recommended in Regression Modeling Strategies to have 5 knots for large data sets 
* estimates less parameters than penalized spline smoothing with AIC criteria 
* **AIC is higher for these splines**

```{r}
library(rms)
model <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + text_all_lower + update + prior_effort + ampm + weekday + rcs(device_length, 5) + sqrt(avg_tag_score) + rcs(text_length, 3) + rcs(avg_tag_length, 3) + rcs(newline_ratio, 3), data = train1)
AIC(model, k = 2) # 64977.65
# with cph fits, keep getting errors with predict function 
```

##### Interactions

* probably not going to include any interactions - don't do much for the model

```{r, echo = FALSE}
# adding interaction between title_questionmark + title_beginwh 
reduced <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + title_beginwh + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday, data = train1) # title_beginwh was not included in result from stepAIC, added in this model to compare it to the model with the interaction

interaction <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + title_beginwh + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + title_questionmark*title_beginwh, data = train1)
summary(interaction)
AIC(interaction, k = 2) # 64963.33 barely drops 
oshitar::compare_nested(interaction, reduced)
```

### Final model(??)

* Regression Modeling Strategies by Frank Harrell pointed out that full models perform the same/if not better than models from stepwise selection
* this model:
    + contains all variables
    + puts splines on all quantitative variables (except for avg_tag_score)
    + device_length might be a better predictor- set a spline with 5 knots, other variables not throught to be as signficant have 3 knots

```{r}
full <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + pspline(text_length, 3) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3), data = train1)
AIC(full, k = 2)
```

### Performance metric on training data set

```{r}
# working with model_spline2 (model with all splines)
train1$predictions <- predict(object = full , type = "risk")
performance <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE)
performance

survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train1) # Concordance = 0.6382697
Hmisc::rcorrcens(Surv(time_until_answer, answered) ~ predictions, data = train1)
```

### Predicting on the test data

```{r, include = FALSE}
# function to perform cross validation
# vars is a string of variables, ex: "new_category + new_user..."
# outputs df with performance metrics (for model on train and test data 
crossval <- function(vars, train, test) {
  formula <- paste("Surv(time_until_answer, answered) ~ ", vars, sep = "")
  model <- survival::coxph(as.formula(formula), data = train)

  train[["predictions"]] <- predict(model, type = "risk")
  metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train)
  train_metrics <- data.frame(HR = exp(metric$coefficients), 
                              LR = round(metric$stats[3],2), 
                              pval = round(metric$stats[5],2), 
                              AIC = stats::AIC(metric, k = 2),
                              Dxy = round(metric$stats[9],2), 
                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train)$concordance)

  # predicting on test data
  test[["predictions"]] <- predict(model, newdata = test, type = "risk")

  # computing performance metrics 
  metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
  test_metrics <- data.frame(HR = exp(metric1$coefficients), 
                             LR = round(metric1$stats[3],2), 
                             pval = round(metric1$stats[5],2), 
                             AIC = stats::AIC(metric1, k = 2),
                             Dxy = round(metric1$stats[9],2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)

  # returns data frame with train/test metrics 
  statistics <- rbind(train_metrics, test_metrics)
  rownames(statistics) <- c("Training Data", "Test Data")
  return(statistics)

}

# function to get average performance metric for each iteration of cv 
# takes list of performance metrics (output from purrr::map(..crossval))
get_avgmetrics <- function(list) {
  avg <- rbind(train_avg = colMeans(purrr::map_df(1:length(list), ~rbind(list[[.]][1,]))), test_avg = colMeans(purrr::map_df(1:length(list), ~rbind(list[[.]][2,]))))
  return(avg)
}

```

```{r, echo = FALSE}
trains <- list(train1, train2, train3, train4, train5)
tests <- list(test1, test2, test3, test4, test5)

# model with penalized AIC splines on device, text, tag length 
allvars <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = 'AIC') + pspline(text_length, df = 0, method = 'AIC') + pspline(avg_tag_length, df = 0, method = 'AIC')"
devtexttag <- purrr::map2(trains, tests, ~crossval(allvars, .x, .y))
devtexttag
# 5th iteration: pvalue for loglikelihood ratio test = 0.51 on test data 
```

```{r}
# penalized AIC spline on device length only 
devvars <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = 'AIC')"
dev <- purrr::map2(trains, tests, ~crossval(devvars, .x, .y))
dev
```

```{r}
# penalized AIC spline on text length only
varstext <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(text_length, df = 0, method = 'AIC')"

text <- purrr::map2(trains, tests, ~crossval(varstext, .x, .y))
```

```{r}
# penalized AIc spline on tag length only 
varstag <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(avg_tag_length, df = 0, method = 'AIC')"

tag <- purrr::map2(trains, tests, ~crossval(varstag, .x, .y))
```

```{r}
# penalized AIc spline on newline ratio only 
varsnew <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(newline_ratio, df = 0, method = 'AIC')"

newline <- purrr::map2(trains, tests, ~crossval(varsnew, .x, .y))
```

```{r}
# penalized splines on dev and text length (both had most signficant p values out of all splines in univariate analysis)
devtextvars <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = 'AIC') + pspline(text_length, df = 0, method = 'AIC')"

devtext <- purrr::map2(trains, tests, ~crossval(devtextvars, .x, .y))

```

```{r}
# with log10(text_length) and spline on dev length (going off of univariate p-values) 
logtextvars <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + pspline(device_length, df = 0, method = 'AIC')"

logtext <- purrr::map2(trains, tests, ~crossval(logtextvars, .x, .y))
```

```{r}
# no splines
vars <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday"
nospline <- purrr::map2(trains, tests, ~crossval(vars, .x, .y))
```

```{r}
# log10(text_length) + sqrt(newline_ratio) + splines: device_length, avg_tag_length + interaction between titlebeginwh*title_questionmark
vars1 <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = 'AIC') + pspline(avg_tag_length, df = 0, method = 'AIC') + title_beginwh + title_questionmark*title_beginwh"

lots <- purrr::map2(trains, tests, ~crossval(vars1, .x, .y))
```

```{r}
# log10(text_length) + sqrt(newline_ratio) + splines: device_length, avg_tag_length, newline_ratio + interaction between titlebeginwh*title_questionmark 
vars2 <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = 'AIC') + pspline(avg_tag_length, df = 0, method = 'AIC') + title_beginwh + pspline(newline_ratio, df = 0, method = 'AIC') + title_questionmark*title_beginwh"

lots2 <- purrr::map2(trains, tests, ~crossval(vars2, .x, .y))
```

```{r}
# splines on device_length and avg_tag_length (FINAL MODEL???)
varsdevtag <- "new_category + as.factor(new_user) + 
    contain_unanswered + title_questionmark + sqrt(avg_tag_score) + 
    text_all_lower + update + prior_effort + ampm + weekday + 
    pspline(device_length, df = 0, method = 'AIC') + pspline(avg_tag_length, 
    df = 0, method = 'AIC')"
devtag <- purrr::map2(trains, tests, ~crossval(varsdevtag, .x, .y))
```

##### full model (all variables) with splines on all continuous variables except for avg_tag_score

```{r}
full <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + pspline(text_length, 3) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3), data = train1)

fullvars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + pspline(text_length, 3) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3)"
full <- purrr::map2(trains, tests, ~crossval(fullvars, .x, .y))
full

ugh <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 0, method = 'AIC') + pspline(text_length, df = 0, method = 'AIC') + pspline(avg_tag_length, df = 0, method = 'AIC') + pspline(newline_ratio, df = 0, method = 'AIC'), data = train1)

summary(ugh) 
AIC(ugh, k = 2)
```

##### viewing all metrics

```{r}
# model I came to while trying to correct PH assumption violation 
# stratified on ampm, orthogonal polynomial on text_length (new_category still violates PH)
fullvars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + strata(ampm) + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + poly(text_length, 2) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3)"
fullcv <- purrr::map2(trains, tests, ~crossval(fullvars, .x, .y))

# turn avg_list into a data frame with rownames???
list <- list(dev, text, tag, newline, devtext, logtext, nospline, lots, lots2, devtag, full)
avg_list <- purrr::map(list, ~get_avgmetrics(.))
```

```{r}
# function to get the average difference between performance metrics for training and test data 
get_avgdiff <- function(list) {
  list_diff <- purrr::map(1:length(list), ~diff(as.matrix(list[[.]], lag = 1)))
  df <- plyr::ldply(list_diff, data.frame) %>%
    select(HR, Dxy, Concordance) %>%
    colMeans()
  return(df)
}

# find a way to put rownames on df_diff in a reproducible way 
l_diff <- purrr::map(list, ~get_avgdiff(.))
df_diff <- plyr::ldply(purrr::map(1:length(l_diff), ~data.frame(as.list(l_diff[[.]]))))
df_diff
#least amount of average change found in model with only the spline on newline_ratio 
```

### FINAL MODEL (?)
```{r}
# final <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + 
#     contain_unanswered + title_questionmark + sqrt(avg_tag_score) + 
#     text_all_lower + update + prior_effort + ampm + weekday + 
#     pspline(device_length, df = 0, method = 'AIC') + pspline(avg_tag_length, 
#     df = 0, method = 'AIC'), data = x)
# summary(final)
# AIC(final, k = 2) # 82957.53

full <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + ampm + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + pspline(text_length, 3) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3), data = x)
AIC(full, k = 2) # 82974.66

x$predictions <- predict(full, type = "risk")
performance <- cph(Surv(time_until_answer, answered) ~ predictions, data = x)
AIC(performance, k = 2) # 83243.04

final_metrics <- data.frame(HR = exp(performance$coefficients), 
                              LR = round(performance$stats[3],2), 
                              pval = round(performance$stats[5],2), 
                              AIC = stats::AIC(performance, k = 2),
                              Dxy = round(performance$stats[9],2), 
                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = x)$concordance)
final_metrics
# performance metrics on entire data set are not different from cross validation 
```

```{r}
# plotting predictions against observed values (calibration plot?)
# final_train1 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + 
#     contain_unanswered + title_questionmark + sqrt(avg_tag_score) + 
#     text_all_lower + update + prior_effort + ampm + weekday + 
#     pspline(device_length, df = 0, method = 'AIC') + pspline(avg_tag_length, 
#     df = 0, method = 'AIC'), data = train1)
# train1$predictions <- predict(final_train1, type = "risk")
# 
# ggplot(train1, aes(x = time_until_answer, y = predictions)) + 
#   geom_point() + 
#   ggtitle("Calibration plot")
# # identify the point at around t = 0 with an incorrect prediction
# 
# test1$predictions <- predict(final_train1, newdata = test1, type = "risk")
# ggplot(train1, aes(x = time_until_answer, y = predictions)) + 
#   geom_point()
```

```{r, echo = FALSE}
# function to take in a data set, fit the model, output performance metrics, and allow you to specify times to predict survival probabilities

predict_survprob <- function(dataset, times) {
  # fitting the model 
  data <- oshitar::variable_setup(dataset)
  model <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = data)

  # performance metrics
  data[["predictions"]] <- predict(model, type = "risk")
  p <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = data)
  metrics <- data.frame(HR = exp(p$coefficients), 
                             LR = round(p$stats[3],2), 
                             pval = round(p$stats[5],2), 
                             R2 = round(p$stats[8],2), 
                             Dxy = round(p$stats[9],2), 
                             AIC = stats::AIC(p, k = 2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = data)$concordance)

  # calculating predicted survival probabilities
  probs <- pec::predictSurvProb(model, newdata = data, times)

  # returns a list containing performance metrics, and data frame with predicted survival probabilities
  output <- list(metrics, probs)
  return(output)
}
```

### residuals, PH assumption

```{r, residuals, echo = FALSE}
full <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + n_tags + gratitude + prior_effort + strata(ampm) + weekday + n_images + sqrt(avg_tag_score) + pspline(device_length, df = 5) + poly(text_length, 2) + pspline(avg_tag_length, 3) + pspline(newline_ratio, 3), data = train1)
# If i fit text_length with poly (orthogonal polynomial), text_length doesn't violate ph anymore 
#=============================================
#martingale residuals
# plot for each continuous predictor, assess adequacy of functional form of predictors, or suggest a potential form 
train1$mart <- residuals(full, type = "martingale")
train1$sqrt_score <- sqrt(train1$avg_tag_score)

ggplot(train1, aes(x = sqrt_score, y = mart)) + 
  geom_point() + 
  geom_smooth() + 
  ggtitle(paste("Martingale Residuals for device_length")) + 
  scale_x_continuous("Device_length")
# what do I assess for spline variables?

#=============================================
# deviance residuals 
# check for outliers identify questions poorly predicted by the model

train1$deviance <- residuals(full, type = "deviance")
plotly::ggplotly(ggplot(train1, aes(x = 1:nrow(train1), y = deviance)) + 
  geom_point() + 
  scale_x_continuous("Question index") + 
  scale_y_continuous(breaks = seq(-4, 4, by = 0.5)) + 
  ggtitle("Deviance Residuals"))
# quite a few questions have a deviance residual of larger than 2.5 (use ggplotly to identify)

# looking at points with deviance residuals > |2.5| (suggested in 417 notes)

# only 2 of these outliers were not answered

#=============================================

spline5 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + log10(text_length) + sqrt(newline_ratio) + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
# schoenfeld residuals
schoen <- as.data.frame(residuals(full, type = "schoenfeld"))
# complete event times sorted from shortest to longest
schoen$comp_times <- sort(train1[train1$answered != 0,]$time_until_answer)

plot_schoen <- function(var) {
  ggplot(schoen, aes(x = comp_times, y = schoen[[var]])) + 
    geom_point() + 
    geom_smooth(method = "loess") + 
    ggtitle(paste("Schoenfeld Residuals for", var)) + 
    scale_y_continuous("Residuals")
}
names <- names(schoen)[-ncol(schoen)]
purrr::map(names, plot_schoen)

plotly::ggplotly(plot_schoen("ampmMorning")) # comptime 2020.89 (id = 390866) is pulling it up? 
plotly::ggplotly(plot_schoen("new_categoryApple Product")) # same question
plotly::ggplotly(plot_schoen("ps(text_length)8")) # same question
plotly::ggplotly(plot_schoen("title_questionmarkTRUE")) # same question!!! 
# plots indicate that new_user, ampm and a few others may have violated ph assumption

#=============================================
# formal test of PH assumption
options(scipen=999)
ph_test <- data.frame(predictors = rownames(cox.zph(spline5)$table), cox.zph(spline5)$table) 
rownames(ph_test) = NULL
ph_test <- ph_test %>%
  arrange(p)

# plots of variables that violated PH assumption in cox.zph test
plot_schoen("new_categoryApple Product")
plot_schoen("new_categoryElectronics")
plot_schoen("new_categoryCamera")
plot_schoen("new_categoryGame Console")
plot_schoen("title_questionmarkTRUE")

#=============================================
# score residuals: check for influential observations. ID quesitons who's predictor values highly effect
score <- as.data.frame(residuals(full, type = "score"))

# function to plot score residuals
plot_score <- function(column) {
  ggplot(score, aes(x = 1:nrow(score), y = score[[column]])) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Score residuals for", column)) + 
    scale_y_continuous("Score Residuals") + 
    scale_x_continuous("Question index")
}
columns <- names(score)
purrr::map(columns, plot_score)
# score residuals show that there are quite a few points that may be influential (check with ggplotly)
```