### 2017-08-26 Hours.Rmd
updated hours

> Summer Research Hours Log

Date   | Hours | Time        | Description
------ | ----- | ----------- | -----------------------------------------
6/26/2017 | 1.5 | 2:30-4:00PM | Integrated R and Github, familiarized myself with GitHub
6/27/2017 | 1 | 11:00-12:00PM | Explored the Q&A forum, read and took notes on part of the study posted on GitHub
6/28/2017 | 3.5 | 11:30-3:00PM | Finished reading and taking notes on the study, learned how to upload to Github through R studio, started the stringr data camp course
6/28/2017 | 5 | 5:00-10:00PM | Continued with data camp courses
6/29/2017 | 3 | 7:00-10:00AM | Continued with data camp courses, brainstormed variables to consider in the model
6/29/2017 | 3.5 | 3:00-6:30PM | Responded to issues, redid stringr course and created a cheatsheet
6/30/2017 | 3.25 | 8:15-11:30AM | Redid/completed stringr course, practiced with stringr functions on the data set
6/30/2017 | 3 | 3:30-6:30PM | Ran into problems with stringr functions and the dataset- tried to get them to work, moved onto the next data camp course
6/30/2017 | 2 | 6:30-9:30PM | Writing functions data camp course + cheatsheet
7/1/2017 | 2 | 8:30-10:30AM | Completed writing functions course, started dplyr course 
7/2/2017 | 1 | 9:00-10:00PM | Continued working on the dplyr course + cheatsheet
7/3/2017 | 2.25 | 10:00-12:15PM | Finished dplyr course
7/3/2017 | 1.5 | 4:00-5:30PM | Practiced applying dplyr functions to the dataset, started on ggplot2 course
7/3/2017 | 2 | 6:30-8:30PM | Continued working on ggplot2 course + cheatsheet
7/4/2017 | 1.5 | 2:00-3:30PM | Continued with ggplot2 course
7/5/2017 | 3 | 9:30-12:30PM | Finished ggplot2 course, practiced using ggplot on the dataset, reviewed STAT 417 notes
7/5/2017 | 2.25 | 2:45-5:00PM | Started the text mining course + cheatsheet
7/6/2017 | 1.5 | 7:30-9:00AM | Continued with text mining course
7/6/2017 | 2.75 | 2:15-5:00PM | Finished text mining course + cheatsheet
7/6/2017 | 2.5 | 6:30-9:00PM | Tried to apply text mining techniques to the dataset, ran into RWeka package, uploaded hours timesheet to Github
7/7/2017 | 4.75 | 11:15-5:00PM | Learned and practiced with functions in the survival package on the data set, came up with descriptive summaries, explored the response variable 
7/10/2017 | 1 | 9:00-10:00AM | Started learning how to fit parametric distributions in r
7/10/2017 | 4 | 1:00- 5:00PM | Experimented with calculating MLEs from the data for different parametric distributions, tried to create probability plots to assess the fits for each distribution
7/10/2017 | 3 | 6:30-9:30PM | Continued working with parametric distributions, experimented with fitdistrplus package to fit parametric distributions
7/11/2017 | 4.25 | 10:45-3:00PM | Tried to get the probability plots to work, used techniques learned in data camp courses and survival analysis to explore the data set and get an idea of what variables may be important 
7/11/2017 | 3.75 | 6:00-9:45PM | Continued to explore the data set 
7/12/2017 | 2.33 | 9:40-12:00PM | Survival analysis interpretations
7/12/2017 | 4.5 | 2:30-7:00PM | Summary statistics + survival curves for variables I haven't looked at yet
7/12/2017 | 1 | 8:30-9:30PM | More on survival curves, ran into a problem with my code
7/13/2017 | 3 | 7:40-10:40AM | More data set exploration, reviewed cox regression notes, prepared for meeting 
7/13/2017 | 1 | 3:00-4:00PM | Appointment with librarian
7/17/2017 | 1.5 | 2:00-3:30PM | Experimented with text mining techniques to figure out the 100 most frequently used words, ran into problems with RWeka package
7/18/2017 | 2 | 8:30-10:30AM | Continued working with text mining techniques
7/18/2017 | 3 | 1:00-4:00PM | Worked with Anthony on identifying most frequently used terms in answered/unanswered questions
7/19/2017 | 2 | 9:40-11:40AM | Continued working on frequently used terms
7/19/2017 | 3.5 | 1:00-4:30PM | Frequently used terms
7/20/2017 | 3 | 8:00-12:00PM | Worked on plotting survival curves in ggplot + meeting with Dr. Pileggi
7/20/2017 | 5.5 | 2:00-7:30PM | Frequently used terms, started on literature review, started trying to reorganize summary statistics report
7/21/2017 | 4.5 | 7:00-11:30AM | Re-organized summary statistics
7/21/2017 | 6 | 3:30-9:30PM | Continued to re-organize, wrote functions/rewrote parts of code
7/22/2017 | 2 | 10:00-12:00PM | Re-organzied freq words file
7/22/2017 | 4 | 3:00-7:00PM | Literature reviews
7/24/2017 | 1.5 | 8:30-10:00AM | Finished literature reviews
7/24/2017 | 3 | 1:00-4:00PM | Explored tags, created new variables
7/24/2017 | 3 | 6:00-9:00PM | Tried to create tag frequency variable, ran into problems, looked into creating own r package
7/25/2017 | 2 | 9:00-11:00AM | Continued to work analyze tag variable
7/25/2017 | 3 | 1:00-4:00PM | Wrote the code for tag variables, tried to fix Java problem
7/27/2017 | 7 | 12:00-7:00PM | Met with the Data Specialist to discuss data visualization ideas, finished working with tags, started with textual features of the data
7/28/2017 | 10.5 | 10:30-9:00PM | Continued with textual features
7/29/2017 | 4 | 5:00-9:00PM | Finished working with text, started on the datetime variable 
7/30/2017 | 7 | 2:30-9:30| Finished datetime, finished all exploratory analyses, read articles from meeting with Data Specialist 
7/31/2017 | 2 | 10:00-12:00PM | Organized stuff, started thinking about the model 
7/31/2017 | 2 | 3:00-5:00PM | Text mining and wordclouds on iPhone and Android Phone subcategories 
8/1/2017 | 3.5 | 12:00-3:30PM | Started the process of fitting the model 
8/1/2017 | 3 | 5:30-8:30PM | Organized/added to files
8/2/2017 | 5 | 1:15-6:15PM | Fitted CR models, changed summary_stats files
8/3/2017 | 5.5 | 12:00-5:30 | Added to summary stats, created ratio of newlines to textlength variable, tried to estimate MLEs, ran in to problems
8/3/2017 | 2.75 | 8:00-10:45 | Continued to try out MLEs, looked into Tex and downloaded mendeley, continued fitting the model
8/4/2017 | 5.5 | 3:00-8:30PM | Fit a model similar to one found in literature reviews, compared results
8/7/2017 | 4.5 | 11:00-3:30PM | Continued with comparing model results to Stack overflow. 
8/7/2017 | 3 | 5:45-8:45 | Used AIC, fit model 
8/8/2017 | 2.5 | 9:30-12:00PM | Model fitting research
8/8/2017 | 3 | 1:00-4:00PM | Worked with Anthony on building an R package 
8/9/2017 | 4.25 | 10:45-3:00PM | Predictive model fitting research, reviewed cox regression notes from STAT 417
8/9/2017 | 3 | 9:30-12:30AM | Found predictive model article, researched terminology from the article, found r packages to use
8/10/2017 | 8 | 8:00-4:00PM | Finished putting functions in r package, organized files, data camp course on cross validation, continued with model fitting research
8/10/2017 | 2 | 6:00-8:00PM | Tried to figure out backwards selection function to fit model
8/11/2017 | 5.5 | 10:30-4:00PM | Researched more model fitting functions, recoded category variable
8/11/2017 | 2 | 8:00-10:00PM | Plotted residuals + predicted on test data set
8/15/2017 | 9.5 | 11:00-8:30PM | Cross validation + model fitting 
8/16/2017 | 5 | 10:30-3:30PM | Wrote function for cross validation 
8/16/2017 | 2.5 | 7:00-9:30PM | Fit model to full data set, wrote function to take data, fit model, and get predicted survival probabilities 
8/17/2017 | 3.25 | 9:00-12:15PM | Met with Dr. Pileggi, started researching splines 
8/17/2017 | 1 | 1:30-2:30PM | More spline research
8/21/2017 | 4 | 2:00-6:00PM | Fit splines with univariate analysis, backward/forward selection on model including splines + square root + log transformed variables 
8/22/2017 | 5.5 | 10:00-3:30PM  | Univariate analysis for quadratic terms of variables, fixed setup function, removed devices from frequently answered/answered terms list
8/22/2017 | 3.75 | 6:15-10:00PM | More work with splines + cross validation 
8/23/2017 | 9.5 | 9:00-6:30 | More spline research, cross-validated on models with different splines, compared models, fit final model(?)
8/23/2017 | 1 | 7:45-8:45PM | Plotted residuals, assessed PH assumption, researched effect of violation of PH assumption 
8/24/2017 | 2 | 10:00-12:00AM | Read Regression Modeling Strategies text book
8/24/2017 | 3.5 | 3:00-6:30PM | Continued reading regression modeling strategies book + other literature, tried to apply it to the data
8/24/2017 | 1.5 | 7:30-9:00PM | Residuals, checked PH assumption, tried to address PH assumption violations 
8/25/2017 | 3 | 8:15-11:15AM | Addressed PH assumption violations, meeting with Dr. Pileggi
8/25/2017 | 4.5 | 12:00-4:30PM | Worked with final model, checked residuals/PH, cross validation, fitted to the full data set
8/25/2017 | 3 | 5:30-8:30PM | Ran into problems trying to predict survival probabilities

>> __Total Hours: 1...__