### 2017-08-31 function_demonstration.Rmd
created a markdown that demonstrates how to use functions to predict failure

---
title: "Example use of prediction functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown demonstrates how to use functions fit_model, variable_setup and predict_failure to get failure probability predictions for questions posted on iFixit's Answers forum.

#### Setting up the data

```{r}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

# separating into data frames for model fitting and predictions for demonstration purposes
data_for_fit <- out[1:4000,]
for_predict <- out[4001:nrow(out),]
```

#### Fitting the model

```{r}
model <- oshitar::fit_model(data_for_fit, summary = TRUE)
```

#### Setting up the prediction data

* setting up the variables required for predictions
* call variable_setup with forpredicting = TRUE so the function won't try and set up the time_until_answer variable

```{r}
for_predict <- oshitar::variable_setup(for_predict, forpredicting = TRUE)
str(for_predict)
```

#### Predicting failure probabilities

##### predicting failure probabilities at default times, for each question's linear predictor/strata values (not predicting on any new data)

```{r}
predict <- oshitar::predict_failure(model)
head(predict)
```

##### predicting failure probabilities at default times for each question in the new data set

```{r}
predict1 <- oshitar::predict_failure(model, newdata = for_predict)
head(predict1)
```

* Interpretations: 
    + The probability that question 1 of the new data set recieves an answer before 30 minutes have passed since it was posted is `r round(predict1[1,1], 2)`
    + The probability that question 1 of the new data set recieves an answer before 1000 hours (~42 days) have passed since it was posted is `r round(predict1[1,6], 2)`

##### predicting failure probabilities at user-specified times for each question in the new data set

```{r}
predict2 <- oshitar::predict_failure(model, newdata = for_predict, times = c(1, 7, 30, 500))
head(predict2)
```

* Interpretations:
    + The probability that question 6 of the new data set recieves an answer before 500 hours (~21 days) have passed since it was posted is `r round(predict2[6,4], 2)`
    + The probability that question 1 of the new data set recieves an answer before 500 hours (~21 days) have passed since it was posted is `r round(predict2[1,4], 2)`