### 2017-09-30 article_graphics.Rmd
created a markdown containing graphics to use in article

---
title: "Tables for Journal Article"
author: "Lisa Oshita"
date: "9/30/2017"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))
x <- oshitar::variable_setup(out)

library(ggplot2); library(survival); library(rms); library(magrittr)
```

### Distribution of Answer Times

```{r}
x$Answered <- NA
x$Answered[x$answered == 1] <- "Answered"
x$Answered[x$answered == 0] <- "Unanswered"

ggplot(x, aes(x = time_until_answer, fill = Answered)) + 
  geom_histogram(bins = 20) + 
  scale_x_continuous("Time (hours)") +
  scale_y_continuous("Number of Questions") + 
  ggtitle("Distribution of Answer Times") + 
  theme(legend.title = element_blank()) # edit position of axis labels + title
```

### Table of KM estimate percentiles

```{r}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
KM <- survfit(surv_object ~ 1, conf.type = "log-log")

q <- quantile(KM, c(0.01, 0.05, 0.10, 0.25, 0.30, 0.40, 0.50, 0.55, 0.58, 0.60, 0.64))
quantiles <- q[[1]]
quantilesdf <- data.frame(Percent = names(quantiles), Time = quantiles)
colnames(quantilesdf) <- c("Percent Answered", "Time (Hours)")
quantilesdf

# add caption with example interpretation?: At 0.042 hours or 2.52 minutes, 1 percent of question posted recieved an Answer
# explain the jump in the data? 
```

### KM Curve

```{r}
library(survminer)
ggsurvplot(KM, data = x, 
           risk.table = "abs_pct", risk.table.title = "Total number at risk (percentage at risk)", 
           cumevents = TRUE, 
           conf.int = TRUE, conf.int.style = "step", 
           censor = FALSE,
           surv.median.line = "hv", 
           xlab = "Time (hours)", 
           xlim = c(0, 100), break.x.by = 10,
           ylim = c(0, 1), break.y.by = 0.1,
           surv.plot.height = 1, ggtheme = theme_bw(), 
           tables.height = 0.15, tables.theme = theme_cleantable(), fontsize = 2, na.rm = TRUE,
           title = "Kaplan-Meier Curve for all Questions",
           legend = "none")

# include caption: Dotted line on graph indicates the median answer time (time at which 50% of question recieved an answer)

ggsurvtable(KM, data = x, risk.table.type = "abs_pct",
            ggtheme = theme_bw())
```

### Device Categories

```{r}
dev <- prop.table(table(x$new_category))

devdf <- data.frame(Device = names(dev), Percent = round(as.numeric(unname(dev))*100,2))
devdf <- devdf %>% dplyr::arrange(desc(Percent))
colnames(devdf)[2] <- "Percent of all Questions"
devdf
```