### 2017-09-30 exploratory analysis/summary_stats.Rmd
minor edits

---
title: "Summary Statistics"
output:
  html_document: default
  pdf_document: default
---

```{r, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))

library(dplyr)
x <- out %>% 
  tbl_df() %>%
  filter(langid == "en")

x$time_until_answer <- (x$first_answer_date - x$post_date)/3600
empty <- which(is.na(x$time_until_answer))
for (i in empty) {
  x$time_until_answer[i] <- (x$download_date[i] - x$post_date[i])/3600
}

library(ggplot2); library(survival); library(epiR); library(survminer); library(dplyr)
```

### About the data set:

* `r nrow(x)` observations
* `r ncol(x)` variables

### Exploring the response: time_until_answer

```{r, echo = FALSE}
summary <- summary(x$time_until_answer)
proportion_answered <- sum(x$answered)/nrow(x)
```

* **Shortest answer time:** `r round(min(x$time_until_answer), 2)` hrs
* **Longest answer time (not answered):** `r round(max(x$time_until_answer), 2)` hrs (`r round(max(x$time_until_answer)/24, 2)` days)
* **mean:** `r round(mean(x$time_until_answer), 2)` hrs (`r round(mean(x$time_until_answer)/24, 2)` days)
* **median:** `r round(median(x$time_until_answer), 2)` hrs
* **first quartile:** `r round(summary[2], 2)` hrs
* **third quartile:** `r round(summary[5], 2)` hrs (`r round(summary[5]/24, 2)` days)
* **IQR range:** `r round(summary[5] - summary[2], 2)` hrs (`r round((summary[5] - summary[2])/24, 2)` days)
* `r round(proportion_answered*100, digits = 2)`% of questions were answered (complete)
* Within an hour, `r round((sum(x$time_until_answer <= 1)/nrow(x))*100,2)`% of the questions have been answered
* Within 5 hours, `r round((sum(x$time_until_answer <= 5)/nrow(x))*100,2)`% of the questions have been answered
* Within 1 day, `r round((sum(x$time_until_answer <= 24)/nrow(x))*100,2)`% of the questions have been answered

```{r, echo = FALSE}
ggplot(x, aes(x = time_until_answer)) + 
  geom_histogram(bins = 10) + 
  scale_x_continuous("Time (hours)") +
  ggtitle("Distribution of answer times")
```

### Survival Analysis Statistics

##### Based off of the Kaplan-Meier Estimator 
```{r, include = FALSE}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")

KM <- survfit(surv_object ~ 1, conf.type = "log-log")

s <- summary(KM)
```

* number of complete events/answered questions: `r s$table[4]` (`r nrow(x)` total questions) 
* smallest survival probability: `r round(min(s$surv), 2)`
* largest survival probability: `r round(max(s$surv), 2)`
* mean survival time: `r round(s$table[5], 2)` hrs (`r round(s$table[5]/24, 2)` days)
* SE of the mean: `r round(s$table[6], 2)` hrs
* Median survival time: `r round(s$table[7], 2)` hrs
* 95% confidence interval for the median: (`r round(s$table[8], 2)` hrs,`r round(s$table[9], 2)` hrs)

##### Percentiles

* Table of percentiles:

```{r, percentiles, echo = FALSE}
quant <- quantile(KM, c(0.01, 0.10, 0.25, 0.30, 0.50, 0.64, 0.75))
round(quant[[1]], 3)
```
* Interpretations:
    + By `r round(quant$quantile[1], 2)` hrs 1% of the questions have been answered
    + By `r round(quant$quantile[2], 2)` hrs 10% of the questions have been answered
    + By `r round(quant$quantile[3], 2)` hrs 25% of the questions have been answered
    + By `r round(quant$quantile[4], 2)` hrs 30% of the questions have been answered
    + By `r round(quant$quantile[5], 2)` hrs 50% of the questions have been answered
    + 75th percentile doesn't exist (lowest survival probability: `r round(min(summary(KM)$surv),2)`)
    + Highest quantile that exists: By `r round(quant$quantile[6], 2)` hrs (`r round(quant$quantile[6]/24, 2)` days) 64% of the questions have been answered

##### Kaplan-Meier Survival Curve

```{r, echo = FALSE}
ggsurvplot(KM, data = x, 
           risk.table = "abs_pct", risk.table.title = "Total number at risk (percentage at risk)",
           cumevents = TRUE, 
           censor = FALSE, 
           conf.int = TRUE, conf.int.style = "step", 
           surv.median.line = "hv", 
           xlab = "Time (hours)", 
           surv.plot.height = 1, ggtheme = theme_bw(),
           tables.height = 0.15, tables.theme = theme_cleantable(), fontsize = 3)
```

##### Kaplan-Meier Survival Curve showing up to 100 hours

```{r, echo = FALSE}
library(survminer)
ggsurvplot(KM, data = x, 
           risk.table = "abs_pct", risk.table.title = "Total number at risk (percentage at risk)", 
           cumevents = TRUE, 
           conf.int = TRUE, conf.int.style = "step", 
           censor = FALSE,
           surv.median.line = "hv", 
           xlab = "Time (hours)", 
           xlim = c(0, 100), break.x.by = 10,
           ylim = c(0, 1), break.y.by = 0.1,
           surv.plot.height = 1, ggtheme = theme_bw(), 
           tables.height = 0.15, tables.theme = theme_cleantable(), fontsize = 2, na.rm = TRUE)
```

* KM survival curve shows that the probability a question survives (remains unanswered) past a certain time drops immediately after question is posted.
* Within ~ 100 hours after posting, the survival probability levels out (though it still decreases at a slow rate), and hovers around 0.35 (never hits 0).
* Dashed marks around the lines represent the point-wise 95% confidence intervals
* Vertical dashed line on the plots indicates the median survival time

##### Reversing the survival curve

* At around 100 hours, the probability a question is answered before time t starts to level out at around ~0.63

```{r, echo = FALSE}
reversed <- data.frame(time = s$time, prob_ans = 1-s$surv)
ggplot(reversed, aes(x = time, y = prob_ans)) + 
  geom_line(na.rm = TRUE) + 
  scale_x_continuous("Time (hours)", limits = c(0,100)) + 
  scale_y_continuous("Probability a question is answered before time t") + 
  ggtitle("Probability a question is answered before a certain time vs. time (hours)")
```

##### Kaplan-Meier Type Instantaneous Hazard

```{r, instant hazard, echo = FALSE}
h.hat <- epi.insthaz(KM, conf.level = 0.95) #instantaneous hazard on basis of KM 
round(h.hat[1:10,], 4)
```

* Max KM instant hazard rate: `r round(max(h.hat$est), 2)` at `r round(h.hat$time[which(h.hat$est == max(h.hat$est))], 2)` hrs 
* Min KM instant hazard rate: `r round(min(h.hat$est), 2)`
* Interpretations: 
    + h(t = 0.0083) = 0.0155: Questions that have not been answered 0.0083 (0.498 min) hours after posting, are estimated to be answered 0.015 times within the next 0.0006 hours (0.036 min). 
    + h(t = 0.0164) = 0.9289: Questions that have not been answered 0.0164 (0.984 min) hours after posting, are estimated to be answered 0.9289 times within the next 0.0008 hours (0.048 min).

```{r, instant hazard plots, echo = FALSE}
ggplot(h.hat, aes(x = time, y = est)) + 
  geom_line() + 
  scale_x_continuous("Time (hours)") + 
  scale_y_continuous("Instantaneous Hazard") + 
  ggtitle("KM Type Instantaneous Hazard Plot")

ggplot(h.hat, aes(x = time, y = est)) + 
  geom_line(na.rm = TRUE) + 
  scale_x_continuous("Time (hours)", limits = c(0,500), breaks = seq(0, 500, by = 50)) + 
  scale_y_continuous("Instantaneous Hazard") + 
  ggtitle("KM Type Instantaneous Hazard Plot showing up to 500 hours")
```

* After about 150 hours, the KM instant hazard of a question getting answered drops to 0 and stays there pretty consistently, with a few small spikes here and there
* Right before that 45th hour, hazard spikes up twice to about 1.25

##### Nelson-Aalen Type Instantaneous Hazard Rate

```{r, echo = FALSE}
fit <- data.frame(surv = s$surv, time = s$time, n.event = s$n.event, n.risk = s$n.risk)
fit$h.tilde <- fit$n.event/fit$n.risk

a <- fit %>%
  select(time = time, estimate = h.tilde, num_experience_event = n.event, num_at_risk = n.risk)
a[1:10,]
```

* Max Nelson-Aalen Type Hazard: `r round(max(fit$h.tilde, na.rm = TRUE),5)` at `r round(fit$time[which(fit$h.tilde == max(fit$h.tilde, na.rm = TRUE))],2)` hours
* Min Nelson-Aalen Type Hazard: `r min(fit$h.tilde, na.rm = TRUE)` at `r round(fit$time[which(fit$h.tilde == min(fit$h.tilde, na.rm = TRUE))],2)` hours
* Interpretations:
    + h(t = 0.0083) = 0.000129: Given that a randomly selected question has not been answered by 0.0083 hours (0.498 min), the estimated probability that a question is answered in the next 0.0006 hours is 0.000129.
    + h(t = 0.0164) = 0.000258: Given that a randomly selected question has not been answered by 0.0164 hours (0.984 min), the estiamted probability that a question is answered in the next 0.0008 hours is 0.000258.

```{r, echo = FALSE}
ggplot(fit, aes(x = time, y = h.tilde)) + 
  geom_line() + 
  scale_x_continuous("Time (hours)") + 
  scale_y_continuous("Instantaneous Hazard") + 
  ggtitle("Nelson-Aalen Type Instantaneous Hazard Plot")

ggplot(fit, aes(x = time, y = h.tilde)) + 
  geom_line(na.rm = TRUE) + 
  scale_x_continuous("Time (hours)", limits = c(0,25)) + 
  scale_y_continuous("Instantaneous Hazard") + 
  ggtitle("Nelson-Aalen Type Instantaneous Hazard Plot, showing up to 25 hours")
```

##### Cumulative Hazard

* Nelson-Aalen plot: rate of increase of H(t) starts to decrease/levels out around ~50 hours (indicates that instantaneous hazard is decreasing)

```{r, echo = FALSE}
#Nelson-Aalen estimator of cumulative hazard
fit$H.tilde <- cumsum(fit$h.tilde)

ggplot(fit, aes(x = time, y = H.tilde)) +
  geom_line(na.rm = TRUE) +
  scale_x_continuous("Time (hours)", limits = c(0,1000), breaks = seq(0, 1000, by = 100)) +
  scale_y_continuous("Cumulative Hazard Rate based off of Nelson Aalen Type") + 
  ggtitle("Cumulative Hazard Plot (Nelson-Aalen)")

#KM estimator of cumulative hazard 
fit$H.hat <- -log(fit$surv)

ggplot(fit, aes(x = time, y = H.hat)) + 
  geom_line() + 
  scale_x_continuous("Time (hours)") + 
  scale_y_continuous("Cumulative Hazard Rate based off of KM Estimates") + 
  ggtitle("Cumulative Hazard Plot (KM)")
```