### 2017-11-13 article submission/data/import the data.Rmd
rearranged files

---
title: "Practice"
author: "Shannon Pileggi"
date: "June 23, 2017"
output: html_document
---

#Import data

```{r}
#getwd() identifies the working directory associated with this Git Hub project

#file.path() creates a file path name that concatenates the working directory with the sub folder for the data, and also automatically adjusts the direction of slashes for your operating system
dir <- file.path(getwd(),"data")

#Now we append the name of the data set to the directory
out <- read.csv(file.path(dir, "answers_data.csv"))

summary(out)

```