### 2017-11-13 model fitting/univariate_analysis.Rmd
added code for graphic to include in article

---
title: "Univariate Analysis"
author: "Lisa Oshita"
date: "8/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))
list <- oshitar::variable_setup(out)
x <- list[[1]]

library(survival); library(stringr); library(rebus)
library(dplyr); library(stats); library(ggplot2)
```

### Setting up cross validation paths

```{r, echo = FALSE}
set.seed(444)
plan <- vtreat::kWayCrossValidation(nrow(x), 5, NULL, NULL)
```

### Analysis on training data

```{r, include = FALSE}
train1 <- x[plan[[1]]$train, ]

pvalues <- data.frame(variable = character(), 
                      pvalue = numeric())

# function to get summary of cox model + append pvalue of variable from loglik test to df
#===================================================================
get_pvalue <- function(variable, factor = FALSE) {
  if (factor == FALSE){
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ train1[[variable]], data = train1)
  } else {
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ as.factor(train1[[variable]]), data = train1)
  }
  summary <- summary(coxmodel)
  df <- data.frame(variable = variable, 
                   pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}

# Univariate analyses (select variables with p-value < 0.01)
#===================================================================
pvalues <- get_pvalue("category")
pvalues <- get_pvalue("new_category")
pvalues <- get_pvalue("n_images")
pvalues <- get_pvalue("n_tags", factor = TRUE)
pvalues <- get_pvalue("new_user")
pvalues <- get_pvalue("weekday")
pvalues <- get_pvalue("ampm")
pvalues <- get_pvalue("title_length")
pvalues <- get_pvalue("text_length")
pvalues <- get_pvalue("device_length")
pvalues <- get_pvalue("text_contain_punct")
pvalues <- get_pvalue("title_questionmark")
pvalues <- get_pvalue("title_beginwh")

# if first letter of title is capitalized 
train1$capital_title <- str_detect(as.character(train1$title), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_title")

# if first letter of text is capitalized
train1$capital_text <- str_detect(as.character(train1$text), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_text")

pvalues <- get_pvalue("text_all_lower")
pvalues <- get_pvalue("prior_effort")
pvalues <- get_pvalue("gratitude")
pvalues <- get_pvalue("greeting")
pvalues <- get_pvalue("update")
pvalues <- get_pvalue("newline_ratio")
pvalues <- get_pvalue("avg_tag_length")
pvalues <- get_pvalue("avg_tag_score")
pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")
```

### splines + quadratic/cubic terms + square root/log transformations

* **use pspline and penalized spline smoothing method with AIC criteria**

```{r, echo = FALSE}
# new function to add p-values from these fits to pvalue data frame
get_pvalue_new <- function(variable, fit) {
  summary <- summary(fit)
  df <- data.frame(variable = variable, pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}

#===================================================================
# title_length
spline_title <- coxph(Surv(time_until_answer, answered) ~ pspline(title_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("title_length_spline", fit = spline_title)

sqrt_title <- coxph(Surv(time_until_answer, answered) ~ sqrt(title_length), data = train1)
pvalues <- get_pvalue_new("title_length_sqrt", fit = sqrt_title)

title2 <- coxph(Surv(time_until_answer, answered) ~ title_length + I(title_length^2), data = train1)
pvalues <- get_pvalue_new("quad_title", fit = title2)

title3 <- coxph(Surv(time_until_answer, answered) ~ title_length + I(title_length^2) + I(title_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_title", fit = title3)

#===================================================================
# text_length
spline_text <- coxph(Surv(time_until_answer, answered) ~ pspline(text_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("text_length_spline", fit = spline_text)

sqrt_text <- coxph(Surv(time_until_answer, answered) ~ sqrt(text_length), data = train1)
pvalues <- get_pvalue_new("text_length_sqrt", fit = sqrt_text)

logtext <- coxph(Surv(time_until_answer, answered) ~ log10(text_length), data = train1)
pvalues <- get_pvalue_new("text_length_log10", fit = logtext)

text2 <- coxph(Surv(time_until_answer, answered) ~ text_length + I(text_length^2), data = train1)
pvalues <- get_pvalue_new("quad_text", fit = text2)

text3 <- coxph(Surv(time_until_answer, answered) ~ text_length + I(text_length^2) + I(text_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_text", fit = text3)

#===================================================================
# device_length
spline_dev <- coxph(Surv(time_until_answer, answered) ~ pspline(device_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("device_length_spline", fit = spline_dev)

sqrt_dev <- coxph(Surv(time_until_answer, answered) ~ sqrt(device_length), data = train1)
pvalues <- get_pvalue_new("device_length_sqrt", fit = sqrt_dev)

dev2 <- coxph(Surv(time_until_answer, answered) ~ device_length + I(device_length^2), data = train1)
pvalues <- get_pvalue_new("quad_dev", fit = dev2)

dev3 <- coxph(Surv(time_until_answer, answered) ~ device_length + I(device_length^2) + I(device_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_dev", fit = dev3)

#===================================================================
# newline_ratio
fit7 <- coxph(Surv(time_until_answer, answered) ~ pspline(newline_ratio, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("newline_ratio_spline", fit = fit7)

fit8 <- coxph(Surv(time_until_answer, answered) ~ sqrt(newline_ratio), data = train1)
pvalues <- get_pvalue_new("newline_ratio_sqrt", fit = fit8)

newline2 <- coxph(Surv(time_until_answer, answered) ~ newline_ratio + I(newline_ratio^2), data = train1)
pvalues <- get_pvalue_new("quad_newline", fit = newline2)

newline3 <- coxph(Surv(time_until_answer, answered) ~ newline_ratio + I(newline_ratio^2) + I(newline_ratio^3), data = train1)
pvalues <- get_pvalue_new("cubic_newline", fit = newline3)

#===================================================================
# avg_tag_length
fit9 <- coxph(Surv(time_until_answer, answered) ~ pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("avg_tag_length_spline", fit = fit9)

fit10 <- coxph(Surv(time_until_answer, answered) ~ sqrt(avg_tag_length), data = train1)
pvalues <- get_pvalue_new("avg_tag_length_sqrt", fit = fit10)

tag2 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_length + I(avg_tag_length^2), data = train1)
pvalues <- get_pvalue_new("quad_tag", fit = tag2)

tag3 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_length + I(avg_tag_length^2) + I(avg_tag_length^3), data = train1)
pvalues <- get_pvalue_new("cubic_tag", fit = tag3)

#===================================================================
# avg_tag_score
fit11 <- coxph(Surv(time_until_answer, answered) ~ pspline(avg_tag_score, df = 0, method = "AIC"), data = train1)
pvalues <- get_pvalue_new("avg_tag_score_spline", fit = fit11)

fit12 <- coxph(Surv(time_until_answer, answered) ~ sqrt(avg_tag_score), data = train1)
pvalues <- get_pvalue_new("avg_tag_score_sqrt", fit = fit12)

score2 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_score + I(avg_tag_score^2), data = train1)
pvalues <- get_pvalue_new("quad_score", fit = score2)

score3 <- coxph(Surv(time_until_answer, answered) ~ avg_tag_score + I(avg_tag_score^2) + I(avg_tag_score^3), data = train1)
pvalues <- get_pvalue_new("cubic_score", fit = score3)
```

### examining univariate analyses results

```{r, echo = FALSE}
options(scipen=999)
pvalues <- pvalues %>% arrange(pvalue)
```

### Include in the article

```{r}
train1 <- x[plan[[1]]$train, ]

pvalues <- data.frame(variable = character(), 
                      pvalue = numeric())

# function to get summary of cox model + append pvalue of variable from loglik test to df
# ---------------------------------------------------------------------------------------
get_pvalue <- function(variable, factor = FALSE) {
  if (factor == FALSE){
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ train1[[variable]], data = train1)
  } else {
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ as.factor(train1[[variable]]), data = train1)
  }
  summary <- summary(coxmodel)
  df <- data.frame(variable = variable, 
                   pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}
# ---------------------------------------------------------------------------------------

pvalues <- get_pvalue("new_category")
pvalues <- get_pvalue("new_user")
pvalues <- get_pvalue("title_questionmark")
pvalues <- get_pvalue("text_contain_punct")
pvalues <- get_pvalue("weekday")
pvalues <- get_pvalue("title_length")
pvalues <- get_pvalue("text_length")
pvalues <- get_pvalue("device_length")
pvalues <- get_pvalue("text_all_lower")
pvalues <- get_pvalue("prior_effort")
pvalues <- get_pvalue("update")
pvalues <- get_pvalue("newline_ratio")
pvalues <- get_pvalue("avg_tag_length")
pvalues <- get_pvalue("avg_tag_score")
pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")

options(scipen=999)
pvalues <- pvalues %>% arrange(pvalue)

rms::cph(Surv(time_until_answer, answered) ~ avg_tag_score, data = x)
rms::cph(Surv(time_until_answer, answered) ~ sqrt(avg_tag_score), data = x)
```