### 2017-11-25 article submission/graphics.Rmd
added draft in Sweave format

---
title: "Tables/Graphs for Article"
author: "Lisa Oshita"
date: "11/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
dir <- file.path(getwd(),"data")
out <- read.csv(file.path(dir, "answers_data.csv"))
list <- oshitar::variable_setup(out)
x <- list[[1]]
library(rms); library(survival); library(ggplot2)
```

```{r}
model <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + 
                   contain_unanswered + contain_answered + title_questionmark + 
                   text_contain_punct + text_all_lower + update + prior_effort + weekday +
                   sqrt(avg_tag_score) + rcs(sqrt(avg_tag_length), 3) + rcs(log10(text_length), 3) +
                   rcs(sqrt(device_length), 4) + rcs(sqrt(newline_ratio), 3), 
                   data = x, 
                   x = TRUE, y = TRUE, 
                   surv = TRUE)
model
```

### Distribution of answer times (Figure 1)

```{r}
x$Answered <- NA
x$Answered[x$answered == 1] <- "Answered"
x$Answered[x$answered == 0] <- "Unanswered"

ggplot(x, aes(x = time_until_answer, fill = Answered)) +
  geom_histogram(bins = 15, alpha = 0.5, position = "identity") + 
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 15),
        plot.title = element_text(size = 17),
        legend.position = c(0.8, 0.8),
        legend.text = element_text(size = 13)) + 
  guides(fill = guide_legend(title = NULL)) + 
  scale_x_continuous("Time (hrs)") + 
  scale_y_continuous("Number of Questions") + 
  ggtitle("Distribution of Answer Times")
```

### KM Curve (Figure 2)

```{r}
surv_object <- Surv(x$time_until_answer, x$answered, type = "right")
KM <- survfit(surv_object ~ 1, conf.type = "log-log")

# dashed line indicates the median survival time
ggsurvplot(KM, data = x, 
           risk.table = FALSE, 
           cumevents = FALSE, 
           conf.int = TRUE, conf.int.style = "step", 
           censor = FALSE,
           surv.median.line = "hv", 
           xlab = "Time (hours)", 
           xlim = c(0, 200), break.x.by = 20,
           ylim = c(0, 1), break.y.by = 0.1,
           surv.plot.height = 1, ggtheme = theme_bw(), 
           tables.height = 0.15, tables.theme = theme_cleantable(), fontsize = 2, na.rm = TRUE,
           title = "Kaplan-Meier Curve for all Questions",
           legend = "none")
```

### Table of percentiles (Table 1)

```{r}
q <- quantile(KM, c(0.25, 0.50, 0.55, 0.58, 0.60, 0.64)) # 64% is largest available survival probability
quantiles <- q$quantile
quantilesdf <- data.frame(Percent = names(quantiles), Time = round(quantiles, 2))
colnames(quantilesdf) <- c("Percent Answered", "Time (Hours)")
quantilesdf
```

### Univariate analysis results (Table 2) see univariate_analysis file

```{r}
set.seed(444)
splitPlan <- vtreat::kWayCrossValidation(nrow(x), 5, NULL, NULL)

train1 <- x[splitPlan[[1]]$train, ]; test1 <- x[splitPlan[[1]]$app, ]
train2 <- x[splitPlan[[2]]$train, ]; test2 <- x[splitPlan[[2]]$app, ]
train3 <- x[splitPlan[[3]]$train, ]; test3 <- x[splitPlan[[3]]$app, ]
train4 <- x[splitPlan[[4]]$train, ]; test4 <- x[splitPlan[[4]]$app, ]
train5 <- x[splitPlan[[5]]$train, ]; test5 <- x[splitPlan[[5]]$app, ]

pvalues <- data.frame(variable = character(), 
                      pvalue = numeric())

# function to get summary of cox model + append pvalue of variable from loglik test to df
# ---------------------------------------------------------------------------------------
get_pvalue <- function(variable, factor = FALSE) {
  if (factor == FALSE){
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ train1[[variable]], data = train1)
  } else {
    coxmodel <- coxph(Surv(time_until_answer, answered) ~ as.factor(train1[[variable]]), data = train1)
  }
  summary <- summary(coxmodel)
  df <- data.frame(variable = variable, 
                   pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}
# ---------------------------------------------------------------------------------------

pvalues <- get_pvalue("new_category")
pvalues <- get_pvalue("new_user")
pvalues <- get_pvalue("title_questionmark")
pvalues <- get_pvalue("text_contain_punct")
pvalues <- get_pvalue("weekday")
pvalues <- get_pvalue("text_length")
pvalues <- get_pvalue("device_length")
train1$logdevice <- log10(train1$device_length)
train1$sqrtdevice <- sqrt(train1$device_length)
pvalues <- get_pvalue("sqrtdevice")
pvalues <- get_pvalue("text_all_lower")
pvalues <- get_pvalue("prior_effort")
pvalues <- get_pvalue("update")
pvalues <- get_pvalue("newline_ratio")
train1$sqrtnewline <- sqrt(train1$newline_ratio)
pvalues <- get_pvalue("sqrtnewline")
train1$sqrt <- sqrt(train1$avg_tag_length)
pvalues <- get_pvalue("sqrt")
pvalues <- get_pvalue("avg_tag_score")
pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")
train1$logtext <- log10(train1$text_length)
pvalues <- get_pvalue("logtext")
options(scipen=999)
pvalues <- pvalues %>% arrange(pvalue)
pvalues

rms::cph(Surv(time_until_answer, answered) ~ avg_tag_score, data = x)
rms::cph(Surv(time_until_answer, answered) ~ sqrt(avg_tag_score), data = x)
rms::cph(Surv(time_until_answer, answered) ~ avg_tag_length, data = train1)

rms::cph(Surv(time_until_answer, answered) ~ sqrt(newline_ratio), data = train1)

ggplot(train1, aes(x = logtext)) + 
  geom_histogram()
ggplot(train1, aes(x = text_length)) + geom_histogram()
ggplot(train1, aes(x = log10(device_length))) + geom_histogram()
ggplot(train1, aes(x = sqrt(newline_ratio))) + geom_histogram()
```

### Average CV metrics (Table 3 + 4) 
```{r}
# function for cross validation
crossval <- function(vars, train, test) {
  formula <- paste("Surv(time_until_answer, answered) ~ ", vars, sep = "")
  model <- rms::cph(as.formula(formula), data = train)

  train[["predictions"]] <- exp(predict(model, type = "lp"))
  metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train)
  train_metrics <- data.frame(HR = exp(metric$coefficients), 
                              LR = round(metric$stats[3],2), 
                              pval = round(metric$stats[5],2),
                              R2 = round(metric$stats[8], 2),
                              AIC = stats::AIC(metric, k = 2),
                              Dxy = round(metric$stats[9],2), 
                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train)$concordance)

  # predicting on test data
  test[["predictions"]] <- exp(predict(model, newdata = test, type = "lp"))

  # computing performance metrics 
  metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
  test_metrics <- data.frame(HR = exp(metric1$coefficients), 
                             LR = round(metric1$stats[3],2), 
                             pval = round(metric1$stats[5],2), 
                             R2 = round(metric$stats[8], 2),
                             AIC = stats::AIC(metric1, k = 2),
                             Dxy = round(metric1$stats[9],2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)

  # returns data frame with train/test metrics 
  statistics <- rbind(train_metrics, test_metrics)
  rownames(statistics) <- c("Training Data", "Test Data")
  return(statistics)
}

# function to average performance metrics, takes list of output from crossval
get_avgmetrics <- function(list) {
  avg <- rbind(train_avg = colMeans(purrr::map_df(1:length(list), ~rbind(list[[.]][1,]))), test_avg = colMeans(purrr::map_df(1:length(list), ~rbind(list[[.]][2,]))))
  return(avg)
}

# ------------------------------------------
# without splines
vars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + text_contain_punct + text_all_lower + update + prior_effort + weekday + sqrt(avg_tag_score) + text_length + device_length + sqrt(avg_tag_length) + newline_ratio"

cv_results <- purrr::map2(trains, tests, ~crossval(vars, .x, .y))
avg_results <- get_avgmetrics(cv_results)
avg_results

# ------------------------------------------
# with splines
vars1 <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + text_contain_punct + text_all_lower + update + prior_effort + weekday + sqrt(avg_tag_score) + rcs(log10(text_length), 3) + rcs(sqrt(avg_tag_length), 3) + rcs(sqrt(device_length), 4) + rcs(sqrt(newline_ratio), 3)"

cv_results1 <- purrr::map2(trains, tests, ~crossval(vars1, .x, .y))
avg_results1 <- get_avgmetrics(cv_results1)
avg_results1
```

```{r}
avg_results1
```

### residuals

```{r}
x$mart <- residuals(model, type = "martingale")

ggplot(x, aes(x = avg_tag_score, y = mart)) + 
  geom_point() + 
  geom_smooth() + 
  ggtitle(paste("Martingale Residuals for average_tag_score")) + 
  scale_x_continuous("Average Tag Score (square root)")

plotm <- function(var) {
  ggplot(x, aes(x = x[[var]], y = mart)) + 
    geom_point() + 
    geom_smooth() + 
    ggtitle(paste("Martingale Residuals for", var)) + 
    scale_x_continuous(var)
}

x$sqrt_length <- sqrt(x$avg_tag_length)
x$sqrt_score <- sqrt(x$avg_tag_score)

plotm("sqrt_score")
plotm("text_length")
plotm("device_length")
plotm("sqrt_length")
plotm("newline_ratio")
```

# final model on full data (table 5)

```{r}
# model statistics + parameters
# figure out how to get p-values 
parameterdf <- data.frame(variable = names(model$coefficients),
                          coefficients = round(unname(model$coefficients), 3))
parameterdf

modelstats <- data.frame(statistic = names(model$stats), 
                         value = round(unname(model$stats), 3))
modelstats

# ------------------------------------------
# cross-validation metrics 
x$predictions <- exp(predict(model, type = "lp"))

metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = x)
AIC(metric, k = 2) # 83124.21

final_metrics <- data.frame(HR = exp(metric$coefficients), 
                              LR = round(metric$stats[3],2), 
                              pval = round(metric$stats[5],2),
                              R2 = round(metric$stats[8], 2),
                              AIC = stats::AIC(metric, k = 2),
                              Dxy = round(metric$stats[9],2), 
                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = x)$concordance)
final_metrics

# ------------------------------------------
ph <- cox.zph(model)
ph_test <- data.frame(predictors = rownames(ph$table), ph$table)
rownames(ph_test) = NULL
ph_test <- ph_test %>% arrange(p) 
ph_test
```

```{r}
final_metrics
```