### 2018-05-03 CSM_poster/main.tex
more edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\setbeamertemplate{caption}[numbered]{}

\usepackage{pifont} % for dings

% puts R2 in bold in table 3
\makeatletter
\DeclareRobustCommand\bfseries{%
  \not@math@alphabet\bfseries\mathbf
  \fontseries\bfdefault\selectfont
  \boldmath % <-- added
}
\makeatother

\usepackage{multirow}
\usepackage{color,soul}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in} % A0 width: 46.8in
\setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{Survival Analysis of Questions Posted on the iFixit Answers Forum} % Poster title

\author{Lisa Oshita\footnote{Frost Research Fellow, recipient of the Frost Undergraduate Student Research Award}, Anthony Pileggi, Shannon Pileggi} % Author(s)

\institute{Department of Statistics, California Polytechnic State University} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{0.5ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Overview}

\vspace{0.5ex}

iFixit's online question and answer forum, \textit{Answers}, features over 120,000 user-asked questions related specifically to device repair. Analysis of question response times can reveal factors that affect how quickly questions receive answers, which can lead to suggestions for how users can ask better questions to minimize response times and for how forum design can be improved.

\vspace{2ex}

\textcolor{dblue!70}{\ding{228}} \textcolor{dblue!70}{\textbf{Objective}} Develop a Cox proportional hazards model to predict the survival probability (probability that a question remains unanswered beyond a certain time \textit{t}), of questions on the forum, with the goal of identifying variables significantly associated with response time.

\vspace{0.5ex}

\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Data}

\textcolor{dblue!70}{\ding{228}} 7,760 questions from April 2017 to July 2017.

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} 63.8\% received an answer

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} Shortest response time: 0.5 hours

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} Longest response time: 2,159 hours (90 days)

\begin{figure}
\vspace{2ex}
\includegraphics[width=1\linewidth]{FIG1.pdf}
\caption{Distribution of response times}
\label{fig1}
\end{figure}

Original variables include: device, question title, text, tags, new user status, post date and answer date. Fourteen variables, capturing textual and user information, were derived.

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------

\begin{block}{Methods}

\textcolor{dblue!70}{\ding{228}} \textcolor{dblue!70}{\textbf{Univariate Analysis}}

\vspace{0.5ex}

Used to identify which variables, as well as the optimal form of continuous variables (if transformations and restricted cubic splines should be applied), to include in the model.

\vspace{1ex}

\textcolor{dblue!70}{\ding{228}} \textcolor{dblue!70}{\textbf{Five Fold Cross-Validation}}

\vspace{0.5ex}

In each iteration, the model was built on the training set and used to predict hazard ratios on the test set.

\vspace{1ex}

\textcolor{dblue!70}{\ding{228}} \textcolor{dblue!70}{\textbf{Assess Performance}}

\vspace{0.5ex}

Predicted hazard ratios were entered into separate Cox models as the single predictor. Resulting $R^2$, concordance, Somers' \textit{Dxy}, partial likelihood ratio (PLR) and p-value, were averaged over each iteration and assessed.

\vspace{1ex}

\textcolor{dblue!70}{\ding{228}} \textcolor{dblue!70}{\textbf{Final Model}}

\vspace{0.5ex}

The model was fit to the full data and proportional hazards assumption was assessed.

\vspace{1.5ex}

\end{block}

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\begin{block}{Cross-Validation Results}

Table \ref{table:1} displays average performance metrics achieved in cross-validation, as well as for the model fit to the full data.

\begin{table}[!htbp]
\vspace{3ex}
\begin{tabular}{|r|r|r|r|r|r|r|r|}
  \hline
  & \textbf{HR} & \textbf{LR} & \textbf{p-value} & \textbf{$R^2$} & \textbf{\textit{Dxy}} & \textbf{C} \\
  \hline
  \textbf{Training} & 2.03 & 937.39  & \textless0.0001 & 0.14 & 0.27 & 0.63 \\
  \textbf{Test}     & 1.99 & 220.83  & \textless0.0001 & 0.14 & 0.26 & 0.63 \\
  \textbf{Full}     & 2.03 & 1165.03 & \textless0.0001 & 0.14 & 0.28 & 0.63 \\
  \hline
\end{tabular}
\caption{Performance metrics (HR: Hazard Ratio, LR: Partial Likelihood Ratio, C: Concordance)}
\label{table:1}
\end{table}
\end{block}

\begin{block}{Final Model Statistics}

\textcolor{dblue!70}{\ding{228}} PLR $=$ 1265.29 (p-value \textless0.0001)

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} $R^2 =$ 0.15

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} Somers' \textit{Dxy} $=$ 0.27

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} Several levels of the device categorization variable violated the proportional hazards assumption

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	INTERPRETING HAZARD
%----------------------------------------------------------------------------------------

\begin{block}{Interpreting Hazard}

Controlling for all other predictors, the estimated hazard of receiving an answer is:

\vspace{1ex}

\textcolor{dblue!70}{\ding{228}} 154\% higher (95\% CI (132\%, 179\%)) for Apple products vs. Android/other phones

\vspace{0.75ex}

\textcolor{dblue!70}{\ding{228}} 13\% lower (95\% CI (7\%, 18\%)) for questions posted on the weekend vs. weekday

\vspace{1ex}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------

\begin{block}{Forum Design Suggestions}

\textcolor{dblue!70}{\ding{228}} Rather than allowing users to enter any tag or device name, restrict options to a drop-down list

\vspace{0.5ex}

\textcolor{dblue!70}{\ding{228}} Include more tips to guide users when asking questions

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

%----------------------------------------------------------------------------------------
\begin{block}{Final Model Coefficients}

\begin{table}[!htbp]
% \centering
\begin{tabular}{|L{10cm} L{17.5cm}|c|c|}
  \hline
 \textbf{Variable} & & \textbf{Coefficient (SE)} & \textbf{p-value} \\ \hline
  Device Category & Apple Product & 0.93 (0.048) & \textless0.0001 \\ 
                  & Camera & -0.32 (0.090) & \\ 
                  & Electronics & -0.01 (0.078)  & \\ 
                  & Game Console & 0.24 (0.083)  & \\ 
                  & Home & 0.34 (0.070) &  \\ 
                  & Other & -0.13 (0.056)  & \\ 
                  & PC & 0.28 (0.060) &  \\ 
                  & Tablet & -0.16 (0.081)  & \\ 
                  & Vehicle & 0.40 (0.069)  & \\
                  & Android/Other Phone (reference) & --- & \\ \hline
  \multicolumn{2}{|l|}{Weekend} & -0.13 (0.033)  & \textless0.0001 \\ \hline
  \multicolumn{2}{|l|}{Text contains end punctuation} & 0.03 (0.050) & 0.613 \\ \hline
  \multicolumn{2}{|l|}{Text is in all lower case} & -0.18 (0.064) &  0.006 \\ \hline
  \multicolumn{2}{|L{30cm}|}{Title contains terms considered frequently-used among answered questions} & 0.05 (0.042) &  0.260 \\ \hline
  \multicolumn{2}{|L{30cm}|}{Title contains terms considered frequently-used among unanswered questions} & -0.28 (0.034)  & \textless0.0001 \\ \hline
  \multicolumn{2}{|L{30cm}|}{Title ends in a question mark} & 0.26 (0.033) &  \textless0.0001 \\ \hline
  \multicolumn{2}{|L{30cm}|}{User edited or added to the question's text after posting it} & 0.30 (0.086)  & 0.001 \\ \hline
  \multicolumn{2}{|L{30cm}|}{User was a member for less than one day before posting} & -0.11 (0.036)  & 0.003 \\ \hline
  \multicolumn{2}{|L{30cm}|}{User made an effort to solve the problem prior to asking the question} & -0.07 (0.036)  & 0.045 \\ \hline
  \multicolumn{2}{|L{30cm}|}{Square root of the average tag frequency} & 2.23 (0.720)  & 0.002 \\ \hline
\end{tabular} 
\caption{Coefficients for predictors in the final model. (Continuous predictors fit with restricted cubic splines are not shown)} 
\label{table:4}
\end{table}

\end{block}

\begin{block}{Acknowledgements}

\small{\rmfamily{This research was supported by the Bill and Linda Frost fund of the California Polytechnic State University of San Luis Obispo. We also thank iFixit for providing access to the CQA data.}} \\

\end{block}

%----------------------------------------------------------------------------------------

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again

\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)

% \begin{block}{Acknowledgements}
% 
% \small{\rmfamily{This research was supported by the Bill and Linda Frost fund of the California Polytechnic State University of San Luis Obispo. We also thank iFixit for providing access to the CQA data.}} \\
% 
% \end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2

\end{column} % End of the second column

% \begin{column}{\sepwid}\end{column} % Empty spacer column

%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

% \setbeamercolor{block title}{fg=red,bg=white} % Change the block title color

% \begin{block}{Acknowledgements}
% 
% \small{\rmfamily{This research was supported by the Bill and Linda Frost fund of the California Polytechnic State University of San Luis Obispo. We also thank iFixit for providing access to the CQA data.}} \\
% 
% \end{block}
% 
% \end{column} % End of the third column
% 
\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}